# Processing Seaflow data (only Gradients 1 through 5)


```{r setup, include=FALSE, message=FALSE, warning=FALSE}
## Libraries
library(tidyverse)
library(arrow)
library(readxl)
library(grid)
library(gridExtra)
library(ggrepel)
library(mclust)

## Output directory
here::i_am("01-process-seaflow.Rmd") ## setwd("~/repos/process-data")
outputdir = "output/seaflow" ##outputdir = "~/Dropbox/data/ocean/seaflow/2025-09-15"
dir.create(outputdir, showWarnings = FALSE)

## Helpers
source("01-helpers.R")

## Where the original, particle-level data is
## (Change to your Dropbox/data directory locally.)
## raw_data_dir <- "~/Dropbox/data/ocean/raw-seaflow/particle-data-2024-09-23"
raw_data_dir <- "~/Dropbox/data/ocean/raw-seaflow/particle-data-2025-11-21"  
```

In this document, we will clean and bin the particle-level seaflow datasets
(only Gradients 1 through 5).


## Setup

This is the overall setup for gridding.

```{r}
## Bin resolution for each dimension
fsc_small_bin_count <- 80
pe_bin_count <- 80
chl_small_bin_count <- 80

## Create grid bins.
grid_bins <- list()
grid_bins$fsc_small <- 2^seq(from = log2(0.0003), to = log2(3200),
                             length = fsc_small_bin_count+1)
grid_bins$pe <- 2^seq(from = log2(1),
                      to = log2(3200),
                      length = pe_bin_count+1)
grid_bins$chl_small <- 2^seq(from = log2(1),
                             to = log2(3200),
                             length = chl_small_bin_count+1)

## Organize them into tables
grid_bin_tables = lapply(grid_bins, function(bins_onedim){
  nbin = length(bins_onedim)
  lapply(1:(nbin-1), function(ii){
    c(coord = ii, left_fencepost = bins_onedim[ii], right_fencepost = bins_onedim[ii+2])
  }) %>% bind_rows()
})
grid_bin_long_table = grid_bin_tables %>% bind_rows(.id = "dimname")
grid_filename = paste0("grid.csv")
write.csv(grid_bin_long_table, file = file.path(outputdir, grid_filename), row.names = FALSE)
```


# Binning five Gradients cruises

First, we will carry out the actual binning.

```{r bin, eval = FALSE}
## Focus on the Gradients cruises
cruises = c("SCOPE_16", ## G1
            "MGL1704", ## G2
            "KM1906", ## G3
            "TN397_740", ## G4
            "TN412_130") ## G5

for(cruisename in cruises){
  print(cruisename)

  ## Read particle-level data.
  vct_dir <- str_glue("{raw_data_dir}/{cruisename}/{cruisename}_vct_slim")
  vct_files <- list.files(vct_dir, "\\.parquet$", full.names = TRUE)
  list_of_df = lapply(1:length(vct_files), function(ifile){ read_parquet(vct_files[ifile]) })
  vct <- bind_rows(list_of_df)
  invisible(gc())
  
  ## Do the actual gridding now
  vct["fsc_small_coord"] <- as.integer(cut(vct[["fsc_small"]],
                                           grid_bins[["fsc_small"]],
                                           labels = FALSE,
                                           right = FALSE))
  vct["pe_coord"] <- as.integer(cut(vct[["pe"]],
                                    grid_bins[["pe"]],
                                    labels = FALSE,
                                    right = FALSE))
  vct["chl_small_coord"] <- as.integer(cut(vct[["chl_small"]],
                                           grid_bins[["chl_small"]],
                                           labels = FALSE,
                                           right = FALSE))

  ## Remove beads
  vct = vct %>% filter(pop != "beads")
  
  ## Gridding the data and summing the Qcs in each 3d grid bin.
  gridded <- vct %>%
    group_by(date, across(ends_with("_coord")), pop) %>%
    summarise(Qc = sum(Qc), .groups="drop") %>%
    select(date, fsc_small_coord, chl_small_coord, pe_coord,  Qc) %>%
    arrange(date, fsc_small_coord, chl_small_coord, pe_coord)

  ## Aggregate to hourly level
  hourly_gridded <- gridded %>%
    group_by(date=floor_date(date, "1 hours"), across(ends_with("_coord"))) %>%
    summarise(Qc=sum(Qc), .groups="drop") %>%
    arrange(date, fsc_small_coord, chl_small_coord, pe_coord)
  invisible(gc())
  
  ## Each cruise's binned data is now saved to a file
  filename = paste0(cruisename, "-gridded-before-cleaning-new.csv")
  outfile = file.path(outputdir, filename)
  write.csv(hourly_gridded, file = outfile, row.names = FALSE)
  cat("wrote to", outfile, fill=TRUE)
}
```

Let's see, for every cruise, 1d plots for all 3 dimensions.

```{r example-viz, fig.width = 12, fig.height = 16}
cruises = c("SCOPE_16", ## G1
            "MGL1704", ## G2
            "KM1906", ## G3
            "TN397_740", ## G4
            "TN412_130") ## G5

for(cruisename in cruises){
  print(cruisename)

  ## Load data
  csv_filename = paste0(cruisename, "-gridded-before-cleaning-new.csv")
  datlist <- csv_to_datlist(outputdir = outputdir,
                            csv_filename = csv_filename)
  ylist <- datlist$ylist
  biomass_list <- datlist$biomass_list
  dates <- datlist$datetimes
  rm(datlist)
  gc()
  
  plotlist = lapply(1:3, function(idim){
  datobj_1d = flowmix::collapse_3d_to_1d(ylist = ylist, countslist = biomass_list, idim = idim)
  flowtrend::plot_1d(ylist = datobj_1d$ylist,
                     countslist = datobj_1d$countslist,  bin = TRUE,
                     x = dates) +
    scale_fill_gradientn(colours = c("white", "black", "yellow", "red")) +
    ggtitle(cruisename) +
    ylab(c("fsc (forward scatter)", "chl", "pe")[idim])
  })
  cowplot::plot_grid(plotlist = plotlist, ncol = 1) %>% print()
}
```


## Data cleaning notes

Here are some points that came up while cleaning data.

1. Gradients 3 has small-diameter particles in many time points that need to be
   manually removed. We'll do this by taking the upper-most bins in diameter and
   removing them altogether.
2. We use a bigger number (80) for the number of bins in fsc because, across all
   cruises, fsc values are more widely variable.
3. fsc stands for forward scatter, which can be converted (using MIE theory) to
   diameter, using Francois Ribalet's algorithm.
4. G5 (`TN412_130`) PE values all fall in the bottom bin. Although at first
   glance it may look like censored points, they aren't! We don't and shouldn't
   remove these! 
5. TODO: G5 (`TN412_130`) requires a bit more work. Some cytograms need to be
   visualized and removed manually.

## Cleaning Gradients 3 


```{r, clean-g3, fig.width = 10, fig.height = 10}
## Load binned data
cruisename = "KM1906"
datadir = "/Users/sangwonh/repos/process-data/output/seaflow/"
csv_filename = paste0(cruisename, "-gridded-before-cleaning.csv")
datlist <- csv_to_datlist(outputdir = "~/repos/process-data/output/seaflow",
                          csv_filename = csv_filename)
ylist <- datlist$ylist
biomass_list <- datlist$biomass_list
dates <- datlist$datetimes

## Plotting before processing
## datobj = flowmix::collapse_3d_to_1d(ylist = ylist, countslist = biomass_list, idim = 1)
## gg_before = my_plotter(datobj$ylist, datobj$countslist, dates, paste0(cruisename, " before any processing"))
gg_init = my_marginal_plotter(ylist, biomass_list, dates)


## New approach (for G3 remove max chl, then bottom fsc_small)
res = my_edge_bin_remove(ylist, biomass_list, dimname = "chl_small", min_or_max = "max")
new_ylist = res$ylist
new_biomass_list = res$biomass_list
gg_after_chl = my_marginal_plotter(new_ylist, new_biomass_list, dates)
cowplot::plot_grid(plotlist = list(gg_init,
                                   gg_after_chl), ncol = 2)

res = my_edge_bin_remove(new_ylist, new_biomass_list, dimname = "fsc_small", min_or_max = "min", slack = 1)
newer_ylist = res$ylist
newer_biomass_list = res$biomass_list
gg_after_chl_and_fsc = my_marginal_plotter(newer_ylist, newer_biomass_list, dates)
cowplot::plot_grid(plotlist = list(gg_init,
                                   gg_after_chl,
                                   gg_after_chl_and_fsc), ncol = 3)

## TODO: continue here.( and make the collapse function somewhat faster if possible)
## Specifically the   flowmix::collapse_3d_to_1d()  function.

newest_dates = dates[-c(21, 136:139)]
newest_ylist = newer_ylist[-c(21, 136:139)]
newest_biomass_list = newer_biomass_list[-c(21, 136:139)]
gg_after_date_removal = my_marginal_plotter(newest_ylist, newest_biomass_list, dates = newest_dates)

cowplot::plot_grid(plotlist = list(gg_init,
                                   gg_after_chl,
                                   gg_after_chl_and_fsc,
                                   gg_after_date_removal), ncol = 2)

## Make the 3d plots for those dates that were removed
cowplot::plot_grid(plotlist = flowtrend::plot_3d(ylist = newer_ylist, countslist = newer_biomass_list, tt = 21),
                   flowtrend::plot_3d(ylist = newer_ylist, countslist = newer_biomass_list, tt = 136),
                   flowtrend::plot_3d(ylist = newer_ylist, countslist = newer_biomass_list, tt = 137),
                   flowtrend::plot_3d(ylist = newer_ylist, countslist = newer_biomass_list, tt = 138),
                   flowtrend::plot_3d(ylist = newer_ylist, countslist = newer_biomass_list, tt = 139),
                   ncol = 3)

## Combine these things back into a csv-saveable table.
dates_list <- new_dates %>% as.list()
recombined_list <- purrr::pmap(list(dates_list, new_ylist, new_biomass_list), function(date_str, coords_matrix, qc_vector) {
  coords_tibble <- as_tibble(coords_matrix, .name_repair = "unique")
  colnames(coords_tibble) <- c("fsc_small_coord", "chl_small_coord", "pe_coord") # Rename columns
  coords_tibble %>%
    mutate(date = date_str, Qc = qc_vector)
})
hourly_gridded_reconstructed <- bind_rows(recombined_list) %>%
  select(date, fsc_small_coord, chl_small_coord, pe_coord, Qc)

## Each cruise is now saved to a file
filename = paste0(cruisename, "-gridded-after-cleaning.csv")
outfile = file.path(outputdir, filename)
write.csv(hourly_gridded_reconstructed, file = outfile, row.names = FALSE)
cat("wrote to", outfile, fill=TRUE)
```

```{r,}
## TODO: continue here (do this for the other cruises too.)
## New approach (for G1, remove PE, chl in order)


```






```{r,}


## Find the cutoff's bin number (in fsc_small)
cutoffs = list()
new_countslist = list()
april_16 = which(lubridate::day(dates) == 16)
april_17 = which(lubridate::day(dates) == 17)
april_18 = which(lubridate::day(dates) == 18)
TT = length(biomass_list)

cowplot::plot_grid(plotlist = plotlist, ncol = 1) %>% print()
par(mfrow = c(10,10), mar = c(1,1,1,1))
TT = 100
for(tt in 1:TT){

  ## 
  one_count = datobj$countslist[[tt]]
  one_y = datobj$ylist[[tt]] 

  ## Find cutoff point
  diffs = one_count  %>% diff()
  aa = which(sapply(1:(length(diffs)-1), function(ii){
    (sign(diffs[ii]) != sign(diffs[ii+1])) &
      (diff(diffs)[ii]>0)
  }))
  a = min(aa, na.rm = TRUE)

  ## Temp
  a = max(aa, na.rm = TRUE)
  ## tt = 556
  plot(one_count, type = 'o', main = paste0("tt=", tt, ", ", dates[[tt]]))
  abline(v=a+1)## temp
  ## Q: Why did those three days need special handling? 
  ## end of temp

  ## if(tt %in% c(april_16, april_17, april_18)){
  ##   a = a[a<=9] 
  ## } else {
  ##   a = a[a<=5] 
  ## }
  stopifnot(length(a) <= 1)
  cutoffs[[tt]] = one_y[a]
}

## Next, removing the points with small fsc_small altogether.
## (Also remove the top-most row in fsc_small and chl_small.)
new_biomass_list = biomass_list
new_ylist = ylist
for(tt in 1:TT){
  one_count = biomass_list[[tt]]
  one_y = ylist[[tt]]
  if( length(cutoffs[[tt]]) > 0 ){
    irow_to_remove1 = which(one_y[,"fsc_small_coord"] <= cutoffs[[tt]])
  } else {
    irow_to_remove1 = c()
  }

  ## Removing the top rows
  irow_to_remove2 = which(one_y[,"fsc_small_coord"] == 80)
  irow_to_remove3 = which(one_y[,"chl_small_coord"] == 40)
  irow_to_remove = c(irow_to_remove1, irow_to_remove2, irow_to_remove3)
  if(length(irow_to_remove) > 0){
  new_ylist[[tt]] = one_y[-irow_to_remove,]
  new_biomass_list[[tt]] = one_count[-irow_to_remove]
  }
}

## Remove these odd time points
## dates[c(21, 136:139)] ## which correspond to these hours:
## "2019-04-11 19:00:00 UTC"
## "2019-04-16 16:00:00 UTC"
## "2019-04-16 17:00:00 UTC"
## "2019-04-16 18:00:00 UTC"
## "2019-04-16 19:00:00 UTC"
new_dates = dates[-c(21, 136:139)]
new_ylist = new_ylist[-c(21, 136:139)]
new_biomass_list = new_biomass_list[-c(21, 136:139)]
stopifnot(length(new_dates)==length(new_ylist))
stopifnot(length(new_dates)==length(new_biomass_list))

datobj_after_correction = flowmix::collapse_3d_to_1d(ylist = new_ylist, countslist = new_biomass_list, idim = 1)
gg_after = my_plotter(datobj_after_correction$ylist,
                  datobj_after_correction$countslist,
                  new_dates,
                  paste0(cruisename, " after processing"))

gg_combined = cowplot::plot_grid(plotlist = list(gg_before, gg_after), ncol = 1, labels = NULL)
plot(gg_combined)
gg_combined %>% 
  ggsave(filename =  file.path(outputdir, paste0(cruisename, "-processing.png")), width = 10, height = 13)


```


It's another matter entirely how to mark data for censored versus non-censored data
```{r}
## For censoring: just remove top and bottom row and consider it censored.
ylist %>% sapply(max) %>% unique()
ylist %>% sapply(min) %>% unique()
```

## Cleaning other Gradients cruises 

Use a similar pipeline: (1) remove edge points in fsc and chl, (2) visually
remove weird time points.

```{r, clean-g1245, fig.width = 12, fig.height = 16}
cruises = c("SCOPE_16", ## G1
            "MGL1704", ## G2
            ## "KM1906", ## G3 ## Already done in the above code block
            "TN397_740", ## G4
            "TN412_130") ## G5
for(cruisename in cruises){

  ## Read in a data file
  csv_filename = paste0(cruisename, "-gridded-before-cleaning.csv") 
  datlist = csv_to_datlist("~/repos/process-data/output/seaflow",
                           csv_filename)
  ylist = datlist$ylist
  biomass_list = datlist$biomass_list
  dates = datlist$datetimes

  
  ## Collapse into 1d for plotting
  datobj = flowmix::collapse_3d_to_1d(ylist = ylist, countslist = biomass_list, idim = 1)
  
  ## Plotting before processing
  gg_before = my_plotter(datobj$ylist, datobj$countslist, dates, paste0(cruisename, " before any processing"))
  gg_before
  
  ## For G2 G3 and G4 (but not G1) remove *top* fsc_small_coord bin altogether
  new_biomass_list = biomass_list
  new_ylist = ylist
  new_dates = dates
    TT = length(ylist)
    for(tt in 1:TT){
      one_count = biomass_list[[tt]]
      one_y = ylist[[tt]]

      ## Remove top row of FSC
      irow_to_remove2 = which(one_y[,"fsc_small_coord"] == 80)

      ## Remove top row of CHL
      irow_to_remove1 = c()
      if(cruisename == "TN412_130"){
        ## irow_to_remove1 = c()
        irow_to_remove1 = which(one_y[,"chl_small_coord"] == 40)
      } else if (cruisename == "SCOPE_16"){
        irow_to_remove1 = which(one_y[,"chl_small_coord"] >= 39)
      } else {
        irow_to_remove1 = which(one_y[,"chl_small_coord"] == 40)
      }

      irow_to_remove = unique(c(irow_to_remove1, irow_to_remove2))##, irow_to_remove3))
      if(length(irow_to_remove) > 0){
        new_ylist[[tt]] = one_y[-irow_to_remove,]
        new_biomass_list[[tt]] = one_count[-irow_to_remove]
      }
    }
  
  ## Plotting after processing
  datobj = flowmix::collapse_3d_to_1d(ylist = new_ylist, countslist = new_biomass_list, idim = 1)
  gg_after = my_plotter(datobj$ylist, datobj$countslist, dates, paste0(cruisename, " after processing"))
  gg_after
  gg_combined = cowplot::plot_grid(plotlist = list(gg_before, gg_after), ncol = 1, labels = NULL)
  plot(gg_combined)
  gg_combined %>% 
    ggsave(filename =  file.path(outputdir, paste0(cruisename, "-processing.png")), width = 10, height = 13)


  ## Combine these things back into a csv-saveable table.
  dates_list <- new_dates %>% as.list()
  recombined_list <- purrr::pmap(list(dates_list, new_ylist, new_biomass_list), function(date_str, coords_matrix, qc_vector) {
    coords_tibble <- as_tibble(coords_matrix, .name_repair = "unique")
    colnames(coords_tibble) <- c("fsc_small_coord", "chl_small_coord", "pe_coord") # Rename columns
    coords_tibble %>%
      mutate(date = date_str, Qc = qc_vector)
  })
  hourly_gridded_reconstructed <- bind_rows(recombined_list) %>%
    select(date, fsc_small_coord, chl_small_coord, pe_coord, Qc)
  
  ## Each cruise is now saved to a file
  filename = paste0(cruisename, "-gridded-after-cleaning.csv")
  outfile = file.path(outputdir, filename)
  write.csv(hourly_gridded_reconstructed, file = outfile, row.names = FALSE)
  cat("wrote to", outfile, fill=TRUE)
}
```


Let's see all 1d plots AFTER the data cleaning.


```{r example-viz2, fig.width = 40, fig.height = 15}
cruises = c("SCOPE_16", ## G1
            "MGL1704", ## G2
            "KM1906", ## G3
            "TN397_740", ## G4
            "TN412_130") ## G5

for(cruisename in cruises){
  print(cruisename)

  ## Load data
  csv_filename = paste0(cruisename, "-gridded-before-cleaning.csv")
  datlist <- csv_to_datlist(outputdir = outputdir,
                            csv_filename = csv_filename)
  ylist <- datlist$ylist
  biomass_list <- datlist$biomass_list
  dates <- datlist$datetimes
  rm(datlist)
  gc()
  
  before_plotlist = lapply(1:3, function(idim){
  datobj_1d = flowmix::collapse_3d_to_1d(ylist = ylist, countslist = biomass_list, idim = idim)
  flowtrend::plot_1d(ylist = datobj_1d$ylist,
                     ## countslist = datobj_1d$countslist,
                     countslist = datobj_1d$countslist %>% sapply(function(a)a/sum(a)),
                     bin = TRUE,
                     x = dates) +
    scale_fill_gradientn(colours = c("white", "black", "yellow", "red")) +
    ggtitle(cruisename) +
    ylab(c("fsc (forward scatter)", "chl", "pe")[idim])
  })


  ## Load data
  csv_filename = paste0(cruisename, "-gridded-after-cleaning.csv")
  datlist <- csv_to_datlist(outputdir = outputdir,
                            csv_filename = csv_filename)
  ylist <- datlist$ylist
  biomass_list <- datlist$biomass_list
  dates <- datlist$datetimes
  rm(datlist)
  gc()
  
  after_plotlist = lapply(1:3, function(idim){
  datobj_1d = flowmix::collapse_3d_to_1d(ylist = ylist, countslist = biomass_list, idim = idim)
  flowtrend::plot_1d(ylist = datobj_1d$ylist,
                     ## countslist = datobj_1d$countslist,  bin = TRUE,
                     countslist = datobj_1d$countslist %>% sapply(function(a)a/sum(a)),
                     bin=TRUE,
                     x = dates) +
    scale_fill_gradientn(colours = c("white", "black", "yellow", "red")) +
    ggtitle(cruisename) +
    ylab(c("fsc (forward scatter)", "chl", "pe")[idim])
  })

  ## cowplot::plot_grid(plotlist = before_plotlist, ncol = 1),
  comparison_plot = cowplot::plot_grid(plotlist = c(before_plotlist, after_plotlist), ncol = 2, byrow = FALSE)
  print(comparison_plot)
  comparison_plot %>% 
    ggsave(filename =  file.path(outputdir, paste0(cruisename, "-after-processing.png")), width = 40, height = 15) 
}
```


## Code to visualize entire 3d cytograms.

Will delete eventually.

```{r, fig.width = 15, fig.height = 15, eval = FALSE}
## Take one cruise
cruisename = "TN412_130"
csv_filename = paste0(cruisename, "-gridded-before-cleaning.csv")
datlist <- csv_to_datlist(outputdir = outputdir,
                          csv_filename = csv_filename)
pre_ylist <- datlist$ylist
pre_biomass_list <- datlist$biomass_list
pre_dates <- datlist$datetimes
rm(datlist)
gc()


## The same cruise after cleaning
cruisename = "TN412_130"
csv_filename = paste0(cruisename, "-gridded-after-cleaning.csv")
datlist <- csv_to_datlist(outputdir = outputdir,
                          csv_filename = csv_filename)
ylist <- datlist$ylist
biomass_list <- datlist$biomass_list
dates <- datlist$datetimes
rm(datlist)
gc()

## Afterwards
TT = length(ylist)
for(tt in 1:TT){
  tt=1
  myplot = flowtrend::plot_3d(ylist = ylist, countslist = biomass_list, tt = tt)
  myplot
  comparison_plot = cowplot::plot_grid(plotlist = c(before_plotlist, after_plotlist), ncol = 2, byrow = FALSE)
  print(comparison_plot)
  comparison_plot %>% 
    ggsave(filename =  file.path(outputdir,
                                 paste0(cruisename, "-3d-plots"),
                                 paste0(cruisename, "-", tt, ".png")),
           width = 40, height = 15) 
}



## For a moment, let's see the particle-level data 
## cruisename = "SCOPE_16"
cruisename = "KM1906"
vct_dir <- str_glue("{raw_data_dir}/{cruisename}/{cruisename}_vct_slim")
vct_files <- list.files(vct_dir, "\\.parquet$", full.names = TRUE)
list_of_df = lapply(1:length(vct_files), function(ifile){ read_parquet(vct_files[ifile]) })
vct <- bind_rows(list_of_df)
invisible(gc())

## Do the actual gridding now
vct["fsc_small_coord"] <- as.integer(cut(vct[["fsc_small"]],
                                         grid_bins[["fsc_small"]],
                                         labels = FALSE,
                                         right = FALSE))
vct["pe_coord"] <- as.integer(cut(vct[["pe"]],
                                  grid_bins[["pe"]],
                                  labels = FALSE,
                                  right = FALSE))
vct["chl_small_coord"] <- as.integer(cut(vct[["chl_small"]],
                                         grid_bins[["chl_small"]],
                                         labels = FALSE,
                                         right = FALSE))

## Remove beads
vct = vct %>% filter(pop != "beads")

hourly_gridded <- vct %>%
    group_by(date=floor_date(date, "1 hours"), across(-c("date", "pop"))) 
    ## summarise(Qc=sum(Qc), .groups="drop") %>%
    ## arrange(date, fsc_small, chl_small, pe)


a = hourly_gridded %>% group_by(date) %>% group_split() %>% .[[100]]
# 3. Generate the 3D plot
a_small = a %>% sample_frac(0.1)

## I'm realizing that 1 or 40 is not the measure of censoring; it's the largest /existant/ bin in that cruise.
max_fsc = a_small %>% summarize(max(fsc_small_coord)) %>% unlist()
max_chl = a_small %>% summarize(max(chl_small_coord)) %>% unlist()
max_pe  = a_small %>% summarize(max(pe_coord)) %>% unlist()

min_fsc = a_small %>% summarize(min(fsc_small_coord)) %>% unlist()
min_chl = a_small %>% summarize(min(chl_small_coord)) %>% unlist()
min_pe  = a_small %>% summarize(min(pe_coord)) %>% unlist()

a_small = a_small %>%
  mutate(color = ifelse(pe_coord==max_pe |
                        pe_coord == min_pe |
                        chl_small_coord == max_chl |
                        chl_small_coord == min_chl |
                        fsc_small_coord == max_fsc |
                        fsc_small_coord == max_fsc, "red", "blue"))

## plot_ly(x=rnorm(91), y = rnorm(91), z=  rnorm(91), type = "scatter3d", mode = "markers" )
library(plotly)
plot_ly(
  data = a_small,
  x = ~fsc_small, 
  y = ~chl_small, 
  z = ~pe, 
  color = ~color,
  size = ~Qc^(1/3),  # Moved to the top-level size argument with a formula tilde
  type = "scatter3d", 
  mode = "markers",
  marker = list(
    sizeref = 5,   # Fixed the spelling typo here
    sizemode = "diameter",
    opacity = 0.8
  )
) %>% 
  # 3. Configure the axes to use log scales
  layout(
    scene = list(
      xaxis = list(
        type = "log",
        title = "log(fsc_small)"
      ),
      yaxis = list(
        type = "log",
        title = "log(chl_small)"
      ),
      zaxis = list(
        type = "log",
        title = "log(pe)"
      )
    ))

dim(a)
a %>% select(fsc_small_coord, chl_small_coord, pe_coord) %>% unique() %>% dim() 

  ## cowplot::plot_grid(plotlist = 
  ##                      list(flowtrend::plot_3d(ylist = pre_ylist, countslist = pre_biomass_list, tt = 510),
  ##                           flowtrend::plot_3d(ylist = pre_ylist, countslist = pre_biomass_list, tt = 511),
  ##                           flowtrend::plot_3d(ylist = pre_ylist, countslist = pre_biomass_list, tt = 512),
  ##                           flowtrend::plot_3d(ylist = pre_ylist, countslist = pre_biomass_list, tt = 513),
  ##                           flowtrend::plot_3d(ylist = ylist, countslist = biomass_list, tt = 510),
  ##                           flowtrend::plot_3d(ylist = ylist, countslist = biomass_list, tt = 511),
  ##                           flowtrend::plot_3d(ylist = ylist, countslist = biomass_list, tt = 512),
  ##                           flowtrend::plot_3d(ylist = ylist, countslist = biomass_list, tt = 513)), ncol = 2, byrow = FALSE)


  ## cowplot::plot_grid(plotlist = 
  ##                      list(flowtrend::plot_3d(ylist = pre_ylist, countslist = pre_biomass_list, tt = 510),
  ##                           flowtrend::plot_3d(ylist = pre_ylist, countslist = pre_biomass_list, tt = 511),
  ##                           flowtrend::plot_3d(ylist = pre_ylist, countslist = pre_biomass_list, tt = 512),
  ##                           flowtrend::plot_3d(ylist = pre_ylist, countslist = pre_biomass_list, tt = 513),
  ##                           flowtrend::plot_3d(ylist = ylist, countslist = biomass_list, tt = 510),
  ##                           flowtrend::plot_3d(ylist = ylist, countslist = biomass_list, tt = 511),
  ##                           flowtrend::plot_3d(ylist = ylist, countslist = biomass_list, tt = 512),
  ##                           flowtrend::plot_3d(ylist = ylist, countslist = biomass_list, tt = 513)), ncol = 2, byrow = FALSE)
```




```{r, fig.width = 15, fig.height = 15, eval = FALSE}
## Take one cruise
cruisename = "MGL1704"
csv_filename = paste0(cruisename, "-gridded-before-cleaning.csv")
datlist <- csv_to_datlist(outputdir = outputdir,
                          csv_filename = csv_filename)
ylist <- datlist$ylist
biomass_list <- datlist$biomass_list
dates <- datlist$datetimes
rm(datlist)
gc()

tt = 120
flowtrend::plot_3d(ylist = ylist, countslist = biomass_list, tt = tt, bin = TRUE, plot_title = dates[[tt]])
```

Actually, now I need to 



<!-- ```{r} -->
<!-- ## 1d cytogram over time points -->
<!-- datobj_1d = flowmix::collapse_3d_to_1d(ylist = ylist, countslist = biomass_list, idim = 1) -->
<!-- flowtrend::plot_1d(ylist = datobj_1d$ylist, -->
<!--                    countslist = datobj_1d$countslist,  bin = TRUE, -->
<!--                    x = dates) + -->
<!--   scale_fill_gradientn(colours = c("white", "black", "yellow", "red")) + -->
<!--   ggtitle(cruisename) + -->
<!--   ylab("fsc (forward scatter)") -->
<!-- ``` -->


<!-- Fitting a flowtrend model for fun. -->

<!-- (Note: When running flowmix or flowtrend, sometimes it's a bit tricky to get the -->
<!-- right hyperparameters; for example, `rho_init` needs to be very small.) -->
