---
title: "Processing Seaflow data"
author: "Sangwon Hyun"
date: "`r Sys.Date()`" 
execute:
  warning: false
  message: false
---


## Processing Seaflow data (only Gradients 1 through 5)


In this document, we will clean and bin the particle-level seaflow datasets
(only Gradients 1 through 5).


This is the overall setup for binning/gridding.

```{r}
#| label: setup
#| include: false
#| message: false
#| warning: false

## Libraries
library(tidyverse)
library(arrow)
library(readxl)
library(grid)
library(gridExtra)
library(ggrepel)
library(mclust)

## Output directory
## here::i_am("01-process-seaflow.Rmd") ## setwd("~/repos/process-data")
## outputdir = "output/seaflow" ##outputdir = "~/Dropbox/data/ocean/seaflow/2025-09-15"
## dir.create(outputdir, showWarnings = FALSE)

## Output directory
## (Change to your Dropbox (e.g., Dropbox/data/ocean/seaflow/.) location locally.)
##outputdir = "~/Dropbox/data/ocean/seaflow" 
outputdir = "~/Dropbox/data/ocean/seaflow-40-bins" 

## Helpers
source(here::here("01-helpers.R"))

## Where the original, particle-level data is
## (Change to your Dropbox/data directory locally.)
## raw_data_dir <- "~/Dropbox/data/ocean/raw-seaflow/particle-data-2024-09-23"
raw_data_dir <- "~/Dropbox/data/ocean/raw-seaflow/particle-data-2025-11-21"
```

Here are some helper functions.


```{r helper-functions}
cat_tab <- function(title, level = 3) {
  # The leading "\n\n" forces a clean break from whatever printed previously
  cat("\n\n", strrep("#", level), " ", title, "\n\n", sep = "")
}

## Saving and embedding the marginal plots
save_and_print_plot <- function(plotlist, cruisename, outputdir, height = 30) {
  combined_plot <- cowplot::plot_grid(plotlist = plotlist, ncol = 1)
  filename <- file.path(outputdir, cruisename, paste0(cruisename, "-censor-removal.png"))
  
  ggsave(filename = filename, plot = combined_plot, width = 10, height = height, limitsize = FALSE)
  cat(sprintf("\n\n![](%s)\n\n", filename))
}

## Writing the reconstructed data to CSV
save_reconstructed_csv <- function(data_df, cruisename, outputdir, suffix) {
  outfile <- file.path(outputdir, cruisename, paste0(cruisename, suffix))
  write.csv(data_df, file = outfile, row.names = FALSE)
  cat("wrote to", outfile, fill = TRUE)
}
```

Here, we set some things up for gridding and binning the particle-level data.

```{r grid-setup}
## Bin resolution for each dimension
fsc_small_bin_count <- 80
##pe_bin_count <- 80
##chl_small_bin_count <- 80
pe_bin_count <- 40
chl_small_bin_count <- 40

## Create grid bins.
grid_bins <- list()
grid_bins$fsc_small <- 2^seq(from = log2(0.0003), to = log2(3200),
                             length = fsc_small_bin_count+1)
grid_bins$pe <- 2^seq(from = log2(1),
                      to = log2(3200),
                      length = pe_bin_count+1)
grid_bins$chl_small <- 2^seq(from = log2(1),
                             to = log2(3200),
                             length = chl_small_bin_count+1)

## Organize them into tables
grid_bin_tables = lapply(grid_bins, function(bins_onedim){
  nbin = length(bins_onedim)
  lapply(1:(nbin-1), function(ii){
    c(coord = ii, left_fencepost = bins_onedim[ii], right_fencepost = bins_onedim[ii+2])
  }) %>% bind_rows()
})
grid_bin_long_table = grid_bin_tables %>% bind_rows(.id = "dimname")
grid_bin_long_table
grid_filename = paste0("grid.csv")
write.csv(grid_bin_long_table, file = file.path(outputdir, grid_filename), row.names = FALSE)

## FYI: if you want to reconstruct grid_bins, use this:
if(FALSE){
  grid_bin_long_table = read.csv(file = file.path(outputdir, grid_filename))
  grid_bins <- reconstruct_grid_bins_from_table(grid_bin_long_table)
}
```



## Binning five Gradients cruises

First, we will carry out the actual binning.

```{r bin, eval = FALSE}
## Focus on the Gradients cruises
cruises = c("SCOPE_16", ## G1
            "MGL1704", ## G2
            "KM1906", ## G3
            "TN397_740", ## G4
            "TN412_130") ## G5

for(cruisename in cruises){

  ## Read particle-level data.
  vct_dir <- str_glue("{raw_data_dir}/{cruisename}/{cruisename}_vct_slim")
  vct_files <- list.files(vct_dir, "\\.parquet$", full.names = TRUE)
  list_of_df = lapply(1:length(vct_files), function(ifile){ read_parquet(vct_files[ifile]) })
  vct <- bind_rows(list_of_df)
  invisible(gc())
  
  ## Do the actual gridding now
  vct["fsc_small_coord"] <- as.integer(cut(vct[["fsc_small"]],
                                           grid_bins[["fsc_small"]],
                                           labels = FALSE,
                                           right = FALSE))
  vct["pe_coord"] <- as.integer(cut(vct[["pe"]],
                                    grid_bins[["pe"]],
                                    labels = FALSE,
                                    right = FALSE))
  vct["chl_small_coord"] <- as.integer(cut(vct[["chl_small"]],
                                           grid_bins[["chl_small"]],
                                           labels = FALSE,
                                           right = FALSE))

  ## Remove beads
  vct = vct %>% filter(pop != "beads")
  
  ## Gridding the data and summing the Qcs in each 3d grid bin.
  gridded <- vct %>%
    group_by(date, across(ends_with("_coord")), pop) %>%
    summarise(Qc = sum(Qc), .groups="drop") %>%
    select(date, fsc_small_coord, chl_small_coord, pe_coord,  Qc) %>%
    arrange(date, fsc_small_coord, chl_small_coord, pe_coord)

  ## Aggregate to hourly level
  hourly_gridded <- gridded %>%
    group_by(date=floor_date(date, "1 hours"), across(ends_with("_coord"))) %>%
    summarise(Qc=sum(Qc), .groups="drop") %>%
    arrange(date, fsc_small_coord, chl_small_coord, pe_coord)
  invisible(gc())
  
  ## Each cruise's binned data is now saved to a file
  filename = paste0(cruisename, "-gridded-before-cleaning.csv")
  dir.create(file.path(outputdir, cruisename))
  outfile = file.path(outputdir, cruisename, filename)
  write.csv(hourly_gridded, file = outfile, row.names = FALSE) 
  cat("wrote to", outfile, fill=TRUE)
}
```


Let's see, for every cruise, 1d plots for all 3 dimensions. This is *before* any
cleaning.



```{r example-viz, fig.width = 12, fig.height = 16}
cruises = c("SCOPE_16", ## G1
            "MGL1704", ## G2
            "KM1906", ## G3
            "TN397_740", ## G4
            "TN412_130") ## G5

for(cruisename in cruises){

  ## Load data
  csv_filename = paste0(cruisename, "-gridded-before-cleaning.csv")
  datlist <- csv_to_datlist(outputdir = file.path(outputdir, cruisename),
                            csv_filename = csv_filename)
  ylist <- datlist$ylist
  biomass_list <- datlist$biomass_list
  dates <- datlist$datetimes
  rm(datlist)
  gc()

  ## Make a 1d forward scatter (fsc) plot over time.
  plotlist = lapply(1:3, function(idim){
    capture.output({
      datobj_1d = flowmix::collapse_3d_to_1d(ylist = ylist, countslist = biomass_list, idim = idim)
    })
    flowtrend::plot_1d(ylist = datobj_1d$ylist,
                       countslist = datobj_1d$countslist,  bin = TRUE,
                       x = dates) +
      scale_fill_gradientn(colours = c("white", "black", "yellow", "red")) +
      ggtitle(cruisename) +
      ylab(c("fsc (forward scatter)", "chl", "pe")[idim])
  })
  cowplot::plot_grid(plotlist = plotlist, ncol = 1) %>% print()

}
```


## Data cleaning notes


Next, we will clean/process the binned data.

Here are some points that came up while cleaning data.

<!-- 1. Gradients 3 has small-diameter particles in many time points that need to be -->
<!--    manually removed. We'll do this by taking the upper-most bins in diameter and -->
<!--    removing them altogether. -->
<!-- 2. We use a bigger number (80) for the number of bins in fsc because, across all -->
<!--    cruises, fsc values are more widely variable. -->
<!-- 3. fsc stands for forward scatter, which can be converted (using MIE theory) to -->
<!--    diameter, using Francois Ribalet's algorithm. -->
<!-- 4. G5 (`TN412_130`) PE values all fall in the bottom bin. Although at first -->
<!--    glance it may look like censored points, they aren't! We don't and shouldn't -->
<!--    remove these!  -->
<!-- 5. TODO: G5 (`TN412_130`) requires a bit more work. Some cytograms need to be -->
<!--    visualized and removed manually. -->

1. The binning was done using 80 bins in FSC dimension, and 40 bins in other
   dimensions. However, FSC in a given cruise usually occupies only 40-ish
   bins. 
2. Most cruises have upper CHL censoring.
3. G5 (`TN412_130`) PE values all fall in the bottom bin. Although at first
   glance these may look like censored points, they aren't! We don't and
   shouldn't remove these! (They are more like *truncated* points.)
4. G1 *might* a level jump in Prochlorococcus levels in the middle of the
   cruise. This is apparent from 1d plots and the series of 3d bin plots.

::: {.panel-tabset}

### Cleaning Gradients 3 

```{r clean-g3}
#| fig-width: 12
#| fig-height: 36 

cruisename = "KM1906"
csv_filename = paste0(cruisename, "-gridded-before-cleaning.csv")
datlist <- csv_to_datlist(outputdir = file.path(outputdir, cruisename), csv_filename = csv_filename)
ylist <- datlist$ylist; biomass_list <- datlist$biomass_list; dates <- datlist$datetimes

gg_init = my_marginal_plotter(ylist, biomass_list, dates)

## Visual tool debugging snapshot
tt = 170
undebug(plotly_particle)
obj = plotly_particle(cruisename, tt, 1,
                      outputdir = "~/Dropbox/data/ocean/seaflow-40-bins",
                      grid_filename = "grid.csv",
                      raw_data_dir = "~/Dropbox/data/ocean/raw-seaflow/particle-data-2025-11-21")
print(obj)

## Remove some specific dates
dates = dates[-c(21, 136:139)]
ylist = ylist[-c(21, 136:139)]
biomass_list = biomass_list[-c(21, 136:139)]
gg_after_date_removal = my_marginal_plotter(ylist = ylist, biomass_list = biomass_list, dates = dates)

## Remove top chl
res = my_edge_bin_remove(ylist = ylist, biomass_list = biomass_list, dimname = "chl_small", min_or_max = "max")
new_ylist = res$ylist; new_biomass_list = res$biomass_list
ylist_cens_after_chl = res$ylist_cens; biomass_list_cens_after_chl = res$biomass_list_cens
gg_after_chl = my_marginal_plotter(new_ylist, new_biomass_list, dates)

## Remove bottom fsc_small (slack=1 means remove bottom 2 bins)
res = my_edge_bin_remove(ylist = new_ylist, biomass_list = new_biomass_list, dimname = "fsc_small", min_or_max = "min", slack = 1)
newer_ylist = res$ylist; newer_biomass_list = res$biomass_list
ylist_cens_after_chl_and_fsc = res$ylist_cens; biomass_list_cens_after_chl_and_fsc = res$biomass_list_cens
gg_after_chl_and_fsc = my_marginal_plotter(ylist = newer_ylist, biomass_list = newer_biomass_list, dates)

## Output visualizations
save_and_print_plot(
  plotlist = list(
    gg_init %>% add_title("Initial", .99),
    gg_after_date_removal %>% add_title("After removing some cytograms", .99),
    gg_after_chl %>% add_title("After CHL removal", .99),
    gg_after_chl_and_fsc %>% add_title("After CHL & FSC removal", .99)
  ),
  cruisename = cruisename, outputdir = outputdir
)

## Save removed cytograms manually for G3
removed_cytogram_plot <- cowplot::plot_grid(plotlist = list(
  flowtrend::plot_3d(ylist = newer_ylist, countslist = newer_biomass_list, tt = 21),
  flowtrend::plot_3d(ylist = newer_ylist, countslist = newer_biomass_list, tt = 136),
  flowtrend::plot_3d(ylist = newer_ylist, countslist = newer_biomass_list, tt = 137),
  flowtrend::plot_3d(ylist = newer_ylist, countslist = newer_biomass_list, tt = 138),
  flowtrend::plot_3d(ylist = newer_ylist, countslist = newer_biomass_list, tt = 139)
), ncol = 1) %>% 
  ggpubr::annotate_figure(top = ggpubr::text_grob(paste0("Hourly cytograms that were removed\n",
                          paste(dates[c(21, 136:139)], collapse = " \n ")), face = "bold", size = 10))

rm_filename = file.path(outputdir, cruisename, paste0(cruisename, "-removed-cytograms.png"))
ggsave(filename = rm_filename, plot = removed_cytogram_plot, width = 8, height = 15)
cat(sprintf("\n\n![](%s)\n\n", rm_filename))

## Save Gridded Data (Regular & Censored)
df_reconstructed <- recombine_to_csv(newer_ylist, newer_biomass_list, dates)
save_reconstructed_csv(df_reconstructed, cruisename, outputdir, "-gridded-after-cleaning.csv")

df_censored <- recombine_censored_data(
  all_ylists = list(ylist_cens_after_chl, ylist_cens_after_chl_and_fsc),
  all_biomass_lists = list(biomass_list_cens_after_chl, biomass_list_cens_after_chl_and_fsc),
  dates = dates
)
save_reconstructed_csv(df_censored, cruisename, outputdir, "-gridded-after-cleaning-censored.csv")
```

### Cleaning Gradients2


```{r clean-g2}
#| fig-width: 12
#| fig-height: 36

cruisename = "MGL1704"
csv_filename = paste0(cruisename, "-gridded-before-cleaning.csv")
datlist <- csv_to_datlist(outputdir = file.path(outputdir, cruisename), csv_filename = csv_filename)
ylist <- datlist$ylist; biomass_list <- datlist$biomass_list; dates <- datlist$datetimes

gg_init = my_marginal_plotter(ylist, biomass_list, dates)

## Visual tool debugging
tt = 296 + 57
obj = plotly_particle(cruisename, tt, 1)
print(obj)

## Remove top chl bin first
res = my_edge_bin_remove(ylist = ylist, biomass_list = biomass_list, dimname = "chl_small", min_or_max = "max")
new_ylist = res$ylist; new_biomass_list = res$biomass_list
ylist_cens_after_chl = res$ylist_cens; biomass_list_cens_after_chl = res$biomass_list_cens
gg_after_chl = my_marginal_plotter(new_ylist, new_biomass_list, dates)

## Remove top fsc
res = my_edge_bin_remove(ylist = new_ylist, biomass_list = new_biomass_list, dimname = "fsc_small", min_or_max = "max")
newer_ylist = res$ylist; newer_biomass_list = res$biomass_list
ylist_cens_after_chl_and_fsc = res$ylist_cens; biomass_list_cens_after_chl_and_fsc = res$biomass_list_cens
gg_after_chl_and_fsc = my_marginal_plotter(newer_ylist, newer_biomass_list, dates)

## Output visualizations
save_and_print_plot(
  plotlist = list(
    gg_init %>% add_title("Initial", .99),
    gg_after_chl %>% add_title("After CHL removal", .99),
    gg_after_chl_and_fsc %>% add_title("After CHL & FSC removal", .99)
  ),
  cruisename = cruisename, outputdir = outputdir
)

## Make more 3d particle plots to get a sense of censoring
ttlist = c(18, 155, 226, 97)
plotlist = lapply(ttlist, function(tt){
  plotly_particle(cruisename, tt, show_censored = TRUE)
})
cowplot::plot_grid(plotlist = plotlist, ncol = 2)

## Save Gridded Data (Regular & Censored)
df_reconstructed <- recombine_to_csv(newer_ylist, newer_biomass_list, dates)
save_reconstructed_csv(df_reconstructed, cruisename, outputdir, "-gridded-after-cleaning.csv")

df_censored <- recombine_censored_data(
  all_ylists = list(ylist_cens_after_chl, ylist_cens_after_chl_and_fsc),
  all_biomass_lists = list(biomass_list_cens_after_chl, biomass_list_cens_after_chl_and_fsc),
  dates = dates
)
save_reconstructed_csv(df_censored, cruisename, outputdir, "-gridded-after-cleaning-censored.csv")
```



### Cleaning Gradients 1  

For G1, remove chl, then pe in order.

```{r clean-g1}

#| fig-width: 12
#| fig-height: 36

cruisename = "SCOPE_16"
csv_filename = paste0(cruisename, "-gridded-before-cleaning.csv")
datlist <- csv_to_datlist(outputdir = file.path(outputdir, cruisename), csv_filename = csv_filename)
ylist <- datlist$ylist; biomass_list <- datlist$biomass_list; dates <- datlist$datetimes

gg_init = my_marginal_plotter(ylist, biomass_list, dates)

## Visual tool debugging
tt = 97
obj = plotly_particle(cruisename, tt, 1)
print(obj)

## Remove top chl
res = my_edge_bin_remove(ylist = ylist, biomass_list = biomass_list, dimname = "chl_small", min_or_max = "max")
new_ylist = res$ylist; new_biomass_list = res$biomass_list
ylist_cens_after_chl = res$ylist_cens; biomass_list_cens_after_chl = res$biomass_list_cens
gg_after_chl = my_marginal_plotter(new_ylist, new_biomass_list, dates)

## Remove bottom PE
res = my_edge_bin_remove(ylist = new_ylist, biomass_list = new_biomass_list, dimname = "pe", min_or_max = "min")
newer_ylist = res$ylist; newer_biomass_list = res$biomass_list
ylist_cens_after_chl_and_pe = res$ylist_cens; biomass_list_cens_after_chl_and_pe = res$biomass_list_cens
gg_after_chl_and_pe = my_marginal_plotter(newer_ylist, newer_biomass_list, dates)

## Output visualizations
save_and_print_plot(
  plotlist = list(
    gg_init %>% add_title("Initial", .99),
    gg_after_chl %>% add_title("After CHL removal", .99),
    gg_after_chl_and_pe %>% add_title("After CHL & PE removal", .99)
  ),
  cruisename = cruisename, outputdir = outputdir
)

## Save Gridded Data (Regular & Censored)
df_reconstructed <- recombine_to_csv(newer_ylist, newer_biomass_list, dates)
save_reconstructed_csv(df_reconstructed, cruisename, outputdir, "-gridded-after-cleaning.csv")

df_censored <- recombine_censored_data(
  all_ylists = list(ylist_cens_after_chl, ylist_cens_after_chl_and_pe),
  all_biomass_lists = list(biomass_list_cens_after_chl, biomass_list_cens_after_chl_and_pe),
  dates = dates
)
save_reconstructed_csv(df_censored, cruisename, outputdir, "-gridded-after-cleaning-censored.csv")
```


### Cleaning Gradients 4


```{r clean-g4}
#| fig-width: 12
#| fig-height: 36

cruisename = "TN397_740"
csv_filename = paste0(cruisename, "-gridded-before-cleaning.csv")
datlist <- csv_to_datlist(outputdir = file.path(outputdir, cruisename), csv_filename = csv_filename)
ylist <- datlist$ylist; biomass_list <- datlist$biomass_list; dates <- datlist$datetimes

gg_init = my_marginal_plotter(ylist, biomass_list, dates)

## Visual tool debugging
tt = 250
obj = plotly_particle(cruisename, tt, 1)
obj

## Remove top chl
res = my_edge_bin_remove(ylist = ylist, biomass_list = biomass_list, dimname = "chl_small", min_or_max = "max")
new_ylist = res$ylist; new_biomass_list = res$biomass_list
ylist_cens_after_chl = res$ylist_cens; biomass_list_cens_after_chl = res$biomass_list_cens
gg_after_chl = my_marginal_plotter(new_ylist, new_biomass_list, dates)

## Remove top FSC 
res = my_edge_bin_remove(ylist = new_ylist, biomass_list = new_biomass_list, dimname = "fsc_small", min_or_max = "max")
newer_ylist = res$ylist; newer_biomass_list = res$biomass_list
ylist_cens_after_chl_and_fsc = res$ylist_cens; biomass_list_cens_after_chl_and_fsc = res$biomass_list_cens
gg_after_chl_and_fsc = my_marginal_plotter(newer_ylist, newer_biomass_list, dates)

## Output visualizations
save_and_print_plot(
  plotlist = list(
    gg_init %>% add_title("Initial", .99),
    gg_after_chl %>% add_title("After CHL removal", .99),
    gg_after_chl_and_fsc %>% add_title("After CHL & FSC removal", .99)
  ),
  cruisename = cruisename, outputdir = outputdir
)

## Save Gridded Data (Regular & Censored)
df_reconstructed <- recombine_to_csv(newer_ylist, newer_biomass_list, dates)
save_reconstructed_csv(df_reconstructed, cruisename, outputdir, "-gridded-after-cleaning.csv")

df_censored <- recombine_censored_data(
  all_ylists = list(ylist_cens_after_chl, ylist_cens_after_chl_and_fsc),
  all_biomass_lists = list(biomass_list_cens_after_chl, biomass_list_cens_after_chl_and_fsc),
  dates = dates
)
save_reconstructed_csv(df_censored, cruisename, outputdir, "-gridded-after-cleaning-censored.csv")
```


### Cleaning Gradients 5

```{r clean-g5}
#| fig-width: 12
#| fig-height: 36

cruisename = "TN412_130" ## G5
csv_filename = paste0(cruisename, "-gridded-before-cleaning.csv")
datlist <- csv_to_datlist(outputdir = file.path(outputdir, cruisename), csv_filename = csv_filename)
ylist <- datlist$ylist; biomass_list <- datlist$biomass_list; dates <- datlist$datetimes

gg_init = my_marginal_plotter(ylist, biomass_list, dates)

## Visual tool debugging
tt = 36  
obj = plotly_particle(cruisename, tt, .1)
print(obj)

## Remove top chl
res = my_edge_bin_remove(ylist = ylist, biomass_list = biomass_list, dimname = "chl_small", min_or_max = "max")
new_ylist = res$ylist; new_biomass_list = res$biomass_list
ylist_cens_after_chl = res$ylist_cens; biomass_list_cens_after_chl = res$biomass_list_cens
gg_after_chl = my_marginal_plotter(new_ylist, new_biomass_list, dates)

## Remove top FSC 
res = my_edge_bin_remove(ylist = new_ylist, biomass_list = new_biomass_list, dimname = "fsc_small", min_or_max = "max")
newer_ylist = res$ylist; newer_biomass_list = res$biomass_list
ylist_cens_after_chl_and_fsc = res$ylist_cens; biomass_list_cens_after_chl_and_fsc = res$biomass_list_cens
gg_after_chl_and_fsc = my_marginal_plotter(newer_ylist, newer_biomass_list, dates)

## Output visualizations
save_and_print_plot(
  plotlist = list(
    gg_init %>% add_title("Initial", .99),
    gg_after_chl %>% add_title("After CHL removal", .99),
    gg_after_chl_and_fsc %>% add_title("After CHL & FSC removal", .99)
  ),
  cruisename = cruisename, outputdir = outputdir
)

## Save Gridded Data (Regular & Censored)
df_reconstructed <- recombine_to_csv(newer_ylist, newer_biomass_list, dates)
save_reconstructed_csv(df_reconstructed, cruisename, outputdir, "-gridded-after-cleaning.csv")

df_censored <- recombine_censored_data(
  all_ylists = list(ylist_cens_after_chl, ylist_cens_after_chl_and_fsc),
  all_biomass_lists = list(biomass_list_cens_after_chl, biomass_list_cens_after_chl_and_fsc),
  dates = dates
)
save_reconstructed_csv(df_censored, cruisename, outputdir, "-gridded-after-cleaning-censored.csv")
```

:::



## Binning the remaining cruises 

Here are the remaining cruises.

```{r the-rest-of-the-cruises, eval = TRUE}
cruisenames_other = c("HOT302", "HOT330", "Thompson_12", "KM1906_towfish",
                      "HOT319", "TN396_740", "Thompson_0", "HOT308", "DeepDOM",
                      "TN267", "HOT331", "GP15_1", "Tokyo_1", "SCOPE_9",
                      "HOT304", "HOT329", "Thompson_9", "SCOPE_17", "HOT317",
                      "KM2206_740", "SCOPE_6", "TN414", "HOT309", "KOK1805",
                      "HOT312", "Thompson_4", "KM1923_751", "HOT335", "SCOPE_3",
                      "HOT299", "HOT332", "SCOPE_11", "HOT324", "MBARI_3",
                      "SCOPE_10", "HOT336", "TN265", "SCOPE_5", "HOT300",
                      "Tokyo_2", "SCOPE_18", "KOK1806", "MBARI_1", "Thompson_6",
                      "MBARI_2", "HOT326", "HOT-294", "HOT325", "TN413",
                      "KM1713", "TN398", "Thompson_3", "SCOPE_19", "SCOPE_1",
                      "HOT337", "SCOPE_2", "KM1712", "HOT301", "HOT318",
                      "HOT303", "CMOP_3", "SCOPE_4", "HOT315", "HOT323",
                      "HOT338", "Thompson_11", "HOT313", "SR1917",
                      "SCOPE-PARAGONII", "SCOPE_Falkor2", "TN266",
                      "KiloMoana_1", "SCOPE-PARAGON", "Thompson_5",
                      "Thompson_1", "SCOPE_13", "Thompson_10", "SCOPE_12",
                      "SCOPE_Falkor1", "KM1920", "KM1919", "KM1923_740",
                      "HOT321", "HOT314", "HOT297", "SCOPE_14", "HOT328",
                      "HOT310", "HOT322", "HOT307", "MESO_SCOPE", "GP15_2",
                      "SCOPE_15")
##"SCOPE_16"#G1
##"MGL1704"#G2
##"KM1906",#G3
##"TN397_740"# G4
##"TN412_130"# G5
```

```{r bin-rest, eval = FALSE}
raw_data_dir <- "~/Dropbox/data/ocean/raw-seaflow/particle-data-2026-09-22"

for(cruisename in cruisenames_other){


  ## Read particle-level data.
  vct_dir <- str_glue("{raw_data_dir}/{cruisename}/{cruisename}_vct_slim")
  vct_files <- list.files(vct_dir, "\\.parquet$", full.names = TRUE)
  list_of_df = lapply(1:length(vct_files), function(ifile){ read_parquet(vct_files[ifile]) })
  vct <- bind_rows(list_of_df)
  invisible(gc())
  
  ## Do the actual gridding now
  vct["fsc_small_coord"] <- as.integer(cut(vct[["fsc_small"]],
                                           grid_bins[["fsc_small"]],
                                           labels = FALSE,
                                           right = FALSE))
  vct["pe_coord"] <- as.integer(cut(vct[["pe"]],
                                    grid_bins[["pe"]],
                                    labels = FALSE,
                                    right = FALSE))
  vct["chl_small_coord"] <- as.integer(cut(vct[["chl_small"]],
                                           grid_bins[["chl_small"]],
                                           labels = FALSE,
                                           right = FALSE))

  ## Remove beads
  vct = vct %>% filter(pop != "beads")
  
  ## Gridding the data and summing the Qcs in each 3d grid bin.
  gridded <- vct %>%
    group_by(date, across(ends_with("_coord")), pop) %>%
    summarise(Qc = sum(Qc), .groups="drop") %>%
    select(date, fsc_small_coord, chl_small_coord, pe_coord,  Qc) %>%
    arrange(date, fsc_small_coord, chl_small_coord, pe_coord)

  ## Aggregate to hourly level
  hourly_gridded <- gridded %>%
    group_by(date=floor_date(date, "1 hours"), across(ends_with("_coord"))) %>%
    summarise(Qc=sum(Qc), .groups="drop") %>%
    arrange(date, fsc_small_coord, chl_small_coord, pe_coord)
  invisible(gc())
  
  ## Each cruise's binned data is now saved to a file
  filename = paste0(cruisename, "-gridded-before-cleaning.csv")
  outfile = file.path(outputdir, cruisename, filename)
  write.csv(hourly_gridded, file = outfile, row.names = FALSE) ## TODO: write into a folder named after the cruise..
  cat("wrote to", outfile, fill=TRUE)
}

invisible(gc())
```



<!-- Some temporary plotting code. -->

<!-- ```{r, eval = FALSE} -->
<!-- for(cruisename in cruisenames_other){ -->
<!--   csv_filename = paste0(cruisename, "-gridded-before-cleaning.csv") -->
<!--   datlist <- csv_to_datlist(outputdir = outputdir, ## TODO:  -->
<!--                             csv_filename = csv_filename) -->
<!--   ylist <- datlist$ylist -->
<!--   biomass_list <- datlist$biomass_list -->
<!--   dates <- datlist$datetimes -->
<!--   rm(datlist) -->
<!--   gc() -->

<!--   ## Make a 1d forward scatter (fsc) plot over time. -->
<!--   plotlist = lapply(1:3, function(idim){ -->
<!--     capture.output({ -->
<!--       datobj_1d = flowmix::collapse_3d_to_1d(ylist = ylist, countslist = biomass_list, idim = idim) -->
<!--     }) -->
<!--     flowtrend::plot_1d(ylist = datobj_1d$ylist, -->
<!--                        countslist = datobj_1d$countslist,  bin = TRUE, -->
<!--                        x = dates) + -->
<!--       scale_fill_gradientn(colours = c("white", "black", "yellow", "red")) + -->
<!--       ggtitle(cruisename) + -->
<!--       ylab(c("fsc (forward scatter)", "chl", "pe")[idim]) -->
<!--   }) -->
<!--   ##cowplot::plot_grid(plotlist = plotlist, ncol = 1) %>% print()  -->

<!--   cowplot::plot_grid(plotlist = plotlist, ncol = 1) %>% -->
<!--     ggsave(file = file.path(outputdir, cruisename, paste0(cruisename, "-1d.png")), width = 15, height = 20) -->
<!-- } -->
<!-- ``` -->



## Processing the remaining cruises

The remaining cruises will be processed more automatically, and with simpler
rules.

1. This is the main cleaning we'll do authomatically -- every top and bottom bin
  in every dimension will be set aside and called censored data (biomass).

2. `Thompson_3` data looks pretty weird. Many cytograms to remove.

3. Next up: many of the cruises have a few cytograms that are weird (and don't
   look like a typical cytogram) and should be removed.

```{r clean-remaining-cruises, eval = FALSE}
#| code-fold: true

for(cruisename in cruisenames_other){

  ## Print the markdown header for the tab
  cat_tab(cruisename)

  ## Load binned data
  csv_filename = paste0(cruisename, "-gridded-before-cleaning.csv")
  datlist <- csv_to_datlist(outputdir = file.path(outputdir, cruisename),
                            csv_filename = csv_filename)
  ylist <- datlist$ylist; biomass_list <- datlist$biomass_list; dates <- datlist$datetimes

  slack_top_fsc = ifelse(cruisename == "Thompson_3", 1, 0)
  normalize_by_total_biomass = TRUE

  ######################################
  ## Process top and bottom edges     ##
  ######################################
  gg_init = my_marginal_plotter(ylist, biomass_list, dates)
  
  res1 = my_edge_bin_remove(ylist = ylist, biomass_list = biomass_list, dimname = "chl_small", min_or_max = "max")
  g1 = my_marginal_plotter(res1$ylist, res1$biomass_list, dates, normalize_by_total_biomass)

  res2 = my_edge_bin_remove(res1, dimname = "chl_small", min_or_max = "min")
  g2 = my_marginal_plotter(res2$ylist, res2$biomass_list, dates, normalize_by_total_biomass)

  res3 = my_edge_bin_remove(res2, dimname = "fsc_small", min_or_max = "max", slack = slack_top_fsc)
  g3 = my_marginal_plotter(res3$ylist, res3$biomass_list, dates, normalize_by_total_biomass)

  res4 = my_edge_bin_remove(res3, dimname = "fsc_small", min_or_max = "min")
  g4 = my_marginal_plotter(res4$ylist, res4$biomass_list, dates, normalize_by_total_biomass)

  res5 = my_edge_bin_remove(res4, dimname = "pe", min_or_max = "max")
  g5 = my_marginal_plotter(res5$ylist, res5$biomass_list, dates, normalize_by_total_biomass)

  res6 = my_edge_bin_remove(res5, dimname = "pe", min_or_max = "min")
  g6 = my_marginal_plotter(res6$ylist, res6$biomass_list, dates, normalize_by_total_biomass)

  ## Save and print unified plot  
  save_and_print_plot(
    plotlist = list(
      gg_init %>% add_title("Initial", .99),
      g1 %>% add_title("After CHL top removal", .99),
      g2 %>% add_title("After CHL bottom removal", .99),
      g3 %>% add_title("After FSC top removal", .99),
      g4 %>% add_title("After FSC bottom removal", .99),
      g5 %>% add_title("After PE top removal", .99),
      g6 %>% add_title("After PE bottom removal", .99)
    ),
    cruisename = cruisename, outputdir = outputdir, height = 70
  )
    
  ###########################################################
  ## Save Gridded Data (Regular & Censored)               ###
  ###########################################################
  df_reconstructed <- recombine_to_csv(res6$ylist, res6$biomass_list, dates)
  save_reconstructed_csv(df_reconstructed, cruisename, outputdir, "-gridded-after-cleaning.csv")

  df_censored <- recombine_censored_data(
      all_ylists = list(res1$ylist_cens, res2$ylist_cens, res3$ylist_cens, 
                        res4$ylist_cens, res5$ylist_cens, res6$ylist_cens),
      all_biomass_lists = list(res1$biomass_list_cens, res2$biomass_list_cens, res3$biomass_list_cens,
                               res4$biomass_list_cens, res5$biomass_list_cens, res6$biomass_list_cens),
      dates = dates
  )
  save_reconstructed_csv(df_censored, cruisename, outputdir, "-gridded-after-cleaning-censored.csv")

  ## Memory Cleanup
  rm(datlist, ylist, biomass_list, dates,
     res1, res2, res3, res4, res5, res6,
     g1, g2, g3, g4, g5, g6, gg_init, 
     df_reconstructed, df_censored)
  
  invisible(gc())
}

## Close the tabset from INSIDE the chunk
cat("\n\n:::\n\n")

```


```{r misc-plotting-code, eval = FALSE}
  ## hourly_gridded_censored_reconstructed  %>% 
  ##   group_by(date, pe_coord) %>%
  ##   summarize(Qc = sum(Qc))  %>% 
  ##   ggplot() +
  ##   geom_raster(aes(x=date, y = pe_coord, fill = Qc)) +
  ##   scale_fill_gradientn(colours = c("white", "black", "yellow", "red"))


  ## hourly_gridded_censored_reconstructed  %>% 
  ##   group_by(date, fsc_small_coord) %>%
  ##   summarize(Qc = sum(Qc))  %>% 
  ##   ggplot() +
  ##   geom_raster(aes(x=date, y = fsc_small_coord, fill = Qc)) +
  ##   scale_fill_gradientn(colours = c("white", "black", "yellow", "red"))
```

::: {.panel-tabset}

```{r clean-remaining-cruises-actual}
#| ref.label: "clean-remaining-cruises"
#| echo: false
#| results: asis
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 84
```




```{r, eval = FALSE}
## Make all 3d plots
for(cruisename in cruisenames_other){

  cruisename = "Thompson_3"
  dir.create(file.path(outputdir, cruisename, "3d-plots"))

  ## Load
  csv_filename = paste0(cruisename, "-gridded-after-cleaning.csv")
  datlist <- csv_to_datlist(outputdir = file.path(outputdir, cruisename), 
                            csv_filename = csv_filename)
  ylist <- datlist$ylist
  biomass_list <- datlist$biomass_list
  dates <- datlist$datetimes
  rm(datlist)
  gc()

  TT = length(ylist)

  for(tt in 1:TT){
    print(tt)

    ## TODO: I fixed collapse_3d_to_2d in flowmix; use this!
    plot_3d_temp(ylist = ylist,
                 biomass_list = biomass_list,
                 tt = tt,
                 dates = dates)  
    ggsave(filename =  file.path(outputdir,
                                 cruisename,
                                 "3d-plots",
                                 paste0(cruisename, "-", tt, ".png")),
           width = 20, height = 7) 
  }
}
```
