---
title: "Processing Seaflow data"
author: "Sangwon Hyun"
date: "`r Sys.Date()`" 
execute:
  warning: false
  message: false
---


## Processing Seaflow data (only Gradients 1 through 5)


In this document, we will clean and bin the particle-level seaflow datasets
(only Gradients 1 through 5).


This is the overall setup for binning/gridding.

```{r}
#| label: setup
#| include: false
#| message: false
#| warning: false
## Libraries
library(tidyverse)
library(arrow)
library(readxl)
library(grid)
library(gridExtra)
library(ggrepel)
library(mclust)

## Output directory
## here::i_am("01-process-seaflow.Rmd") ## setwd("~/repos/process-data")
## outputdir = "output/seaflow" ##outputdir = "~/Dropbox/data/ocean/seaflow/2025-09-15"
## dir.create(outputdir, showWarnings = FALSE)

## Output directory
outputdir = "~/Dropbox/data/ocean/seaflow" 

## Helpers
source(here::here("01-helpers.R"))

## Where the original, particle-level data is
## (Change to your Dropbox/data directory locally.)
# raw_data_dir <- "~/Dropbox/data/ocean/raw-seaflow/particle-data-2024-09-23"
raw_data_dir <- "~/Dropbox/data/ocean/raw-seaflow/particle-data-2025-11-21"

## Bin resolution for each dimension
fsc_small_bin_count <- 80
pe_bin_count <- 80
chl_small_bin_count <- 80

## Create grid bins.
grid_bins <- list()
grid_bins$fsc_small <- 2^seq(from = log2(0.0003), to = log2(3200),
                             length = fsc_small_bin_count+1)
grid_bins$pe <- 2^seq(from = log2(1),
                      to = log2(3200),
                      length = pe_bin_count+1)
grid_bins$chl_small <- 2^seq(from = log2(1),
                             to = log2(3200),
                             length = chl_small_bin_count+1)

## Organize them into tables
grid_bin_tables = lapply(grid_bins, function(bins_onedim){
  nbin = length(bins_onedim)
  lapply(1:(nbin-1), function(ii){
    c(coord = ii, left_fencepost = bins_onedim[ii], right_fencepost = bins_onedim[ii+2])
  }) %>% bind_rows()
})
grid_bin_long_table = grid_bin_tables %>% bind_rows(.id = "dimname")
grid_bin_long_table
grid_filename = paste0("grid.csv")
write.csv(grid_bin_long_table, file = file.path(outputdir, grid_filename), row.names = FALSE)

## FYI: if you want to reconstruct grid_bins, use this:
if(FALSE){
  grid_bin_long_table = read.csv(file = file.path(outputdir, grid_filename))
  grid_bins <- reconstruct_grid_bins_from_table(grid_bin_long_table)
}
```



## Binning five Gradients cruises

First, we will carry out the actual binning.

```{r bin}
#| eval: FALSE
## Focus on the Gradients cruises
cruises = c("SCOPE_16", ## G1
            "MGL1704", ## G2
            "KM1906", ## G3
            "TN397_740", ## G4
            "TN412_130") ## G5

for(cruisename in cruises){

  ## Read particle-level data.
  vct_dir <- str_glue("{raw_data_dir}/{cruisename}/{cruisename}_vct_slim")
  vct_files <- list.files(vct_dir, "\\.parquet$", full.names = TRUE)
  list_of_df = lapply(1:length(vct_files), function(ifile){ read_parquet(vct_files[ifile]) })
  vct <- bind_rows(list_of_df)
  invisible(gc())
  
  ## Do the actual gridding now
  vct["fsc_small_coord"] <- as.integer(cut(vct[["fsc_small"]],
                                           grid_bins[["fsc_small"]],
                                           labels = FALSE,
                                           right = FALSE))
  vct["pe_coord"] <- as.integer(cut(vct[["pe"]],
                                    grid_bins[["pe"]],
                                    labels = FALSE,
                                    right = FALSE))
  vct["chl_small_coord"] <- as.integer(cut(vct[["chl_small"]],
                                           grid_bins[["chl_small"]],
                                           labels = FALSE,
                                           right = FALSE))

  ## Remove beads
  vct = vct %>% filter(pop != "beads")
  
  ## Gridding the data and summing the Qcs in each 3d grid bin.
  gridded <- vct %>%
    group_by(date, across(ends_with("_coord")), pop) %>%
    summarise(Qc = sum(Qc), .groups="drop") %>%
    select(date, fsc_small_coord, chl_small_coord, pe_coord,  Qc) %>%
    arrange(date, fsc_small_coord, chl_small_coord, pe_coord)

  ## Aggregate to hourly level
  hourly_gridded <- gridded %>%
    group_by(date=floor_date(date, "1 hours"), across(ends_with("_coord"))) %>%
    summarise(Qc=sum(Qc), .groups="drop") %>%
    arrange(date, fsc_small_coord, chl_small_coord, pe_coord)
  invisible(gc())
  
  ## Each cruise's binned data is now saved to a file
  filename = paste0(cruisename, "-gridded-before-cleaning.csv")
  outfile = file.path(outputdir, filename)
  write.csv(hourly_gridded, file = outfile, row.names = FALSE) ## TODO: write into a folder named after the cruise..
  cat("wrote to", outfile, fill=TRUE)
}
```


Let's see, for every cruise, 1d plots for all 3 dimensions. This is *before* any
cleaning.



```{r example-viz}
#| output: false
#| fig.width: 12
#| fig.height: 16
cruises = c("SCOPE_16", ## G1
            "MGL1704", ## G2
            "KM1906", ## G3
            "TN397_740", ## G4
            "TN412_130") ## G5

for(cruisename in cruises){

  ## Load data
  csv_filename = paste0(cruisename, "-gridded-before-cleaning.csv")
  datlist <- csv_to_datlist(outputdir = outputdir, ## TODO: 
                            csv_filename = csv_filename)
  ylist <- datlist$ylist
  biomass_list <- datlist$biomass_list
  dates <- datlist$datetimes
  rm(datlist)
  gc()

  ## Make a 1d forward scatter (fsc) plot over time.
  plotlist = lapply(1:3, function(idim){
  datobj_1d = flowmix::collapse_3d_to_1d(ylist = ylist, countslist = biomass_list, idim = idim)
  flowtrend::plot_1d(ylist = datobj_1d$ylist,
                     countslist = datobj_1d$countslist,  bin = TRUE,
                     x = dates) +
    scale_fill_gradientn(colours = c("white", "black", "yellow", "red")) +
    ggtitle(cruisename) +
    ylab(c("fsc (forward scatter)", "chl", "pe")[idim])
  })
  cowplot::plot_grid(plotlist = plotlist, ncol = 1) %>% print()
}
```


## Data cleaning notes

Next, we will clean/process the binned data.

Here are some points that came up while cleaning data.

<!-- 1. Gradients 3 has small-diameter particles in many time points that need to be -->
<!--    manually removed. We'll do this by taking the upper-most bins in diameter and -->
<!--    removing them altogether. -->
<!-- 2. We use a bigger number (80) for the number of bins in fsc because, across all -->
<!--    cruises, fsc values are more widely variable. -->
<!-- 3. fsc stands for forward scatter, which can be converted (using MIE theory) to -->
<!--    diameter, using Francois Ribalet's algorithm. -->
<!-- 4. G5 (`TN412_130`) PE values all fall in the bottom bin. Although at first -->
<!--    glance it may look like censored points, they aren't! We don't and shouldn't -->
<!--    remove these!  -->
<!-- 5. TODO: G5 (`TN412_130`) requires a bit more work. Some cytograms need to be -->
<!--    visualized and removed manually. -->

1. All binning was done using 80 bins in each direction (dimension). FSC in a
   given cruise usually occupies only 40-ish bins. So later, prior to fitting
   flowmix/flowtrend models, we should multiply the FSC bin number by 2.
2. Most cruises have upper CHL censoring.
3. G5 (`TN412_130`) PE values all fall in the bottom bin. Although at first
   glance these may look like censored points, they aren't! We don't and
   shouldn't remove these! (They are more like *truncated* points.)
4. I've found that G1 has a level jump in Prochlorococcus levels in the middle
   of the cruise. This is apparent from 1d plots and the series of 3d bin plots.

### Cleaning Gradients 3 

```{r, clean-g3}
#| fig-width: 12
#| fig-height: 12
#| output: false

## Load binned data
cruisename = "KM1906"
datadir = "~/Dropbox/data/ocean/seaflow/old"
csv_filename = paste0(cruisename, "-gridded-before-cleaning-new.csv")
datlist <- csv_to_datlist("~/Dropbox/data/ocean/seaflow/old",
                          csv_filename = csv_filename)
ylist <- datlist$ylist
biomass_list <- datlist$biomass_list
dates <- datlist$datetimes

## Plotting before processing
gg_init = my_marginal_plotter(ylist, biomass_list, dates)

## Making a 3d scatterplot as a visual tool; (1) you can clearly see censoring
## happening at the top Chl slice/bin, and (2) the bottom two fsc_small bins
## we'll also remove.
tt = 170 ## print(dates[tt])
# obj = plotly_particle(
#   cruisename, tt, outputdir = outputdir, raw_data_dir = raw_data_dir, 
#   show_censored = TRUE
# )
# print(obj)


## Remove top chl first
res = my_edge_bin_remove(ylist, biomass_list, dimname = "chl_small", min_or_max = "max")
new_ylist = res$ylist
new_biomass_list = res$biomass_list
gg_after_chl = my_marginal_plotter(new_ylist, new_biomass_list, dates)


## After removing top chl, then remove bottom fsc_small (slack=1 means remove bottom 2 bins)
res = my_edge_bin_remove(new_ylist, new_biomass_list, dimname = "fsc_small", min_or_max = "min", slack = 1)
newer_ylist = res$ylist
newer_biomass_list = res$biomass_list
gg_after_chl_and_fsc = my_marginal_plotter(newer_ylist, newer_biomass_list, dates)

## Remove some specific dates
newest_dates = dates[-c(21, 136:139)]
newest_ylist = newer_ylist[-c(21, 136:139)]
newest_biomass_list = newer_biomass_list[-c(21, 136:139)]
gg_after_date_removal = my_marginal_plotter(newest_ylist, newest_biomass_list, dates = newest_dates)

# Print the object so Quarto captures it for the HTML output
combined_plot = cowplot::plot_grid(plotlist = list(gg_init %>% add_title("initial",.99),
                                   gg_after_chl %>% add_title("After CHL removal",.99),
                                   gg_after_chl_and_fsc %>% add_title("After CHL&FSC removal",.99), 
                                   gg_after_date_removal%>% add_title("After removing some cytograms",.99)), ncol = 2)  
plot(combined_plot)


# 2. Save the PDF to your directory
ggsave(filename = file.path(outputdir, paste0(cruisename, "-censor-removal.pdf")),
       plot = combined_plot, width = 20, height = 20)


## Make the 3d plots for those dates that were removed
removed_cytogram_plot <-
  cowplot::plot_grid(plotlist =
                       list(flowtrend::plot_3d(ylist = newer_ylist, countslist = newer_biomass_list, tt = 21),
                            flowtrend::plot_3d(ylist = newer_ylist, countslist = newer_biomass_list, tt = 136),
                            flowtrend::plot_3d(ylist = newer_ylist, countslist = newer_biomass_list, tt = 137),
                            flowtrend::plot_3d(ylist = newer_ylist, countslist = newer_biomass_list, tt = 138),
                            flowtrend::plot_3d(ylist = newer_ylist, countslist = newer_biomass_list, tt = 139)),
                     ncol = 1) %>% 
  ggpubr::annotate_figure(top = ggpubr::text_grob(paste0(
                                            "Hourly cytogram that were removed\n",
                                            paste(dates[c(21, 136:139)], collapse = " \n ")),
                                            color = "black", face = "bold", size = 10))

ggsave(filename =  file.path(outputdir,
                             paste0(cruisename, "-removed-cytograms.pdf")),
       plot = removed_cytogram_plot,
       width = 8, height = 15)

## Make all 3d plots
if(FALSE){
  dir.create(file.path(outputdir, paste0(cruisename, "-3d-plots")))
  list.files(outputdir)
  TT = length(newest_ylist)
  for(tt in 1:TT){
    print(tt)
    plot_3d_temp(ylist = newest_ylist,
                 biomass_list = newest_biomass_list,
                 tt = tt,
                 dates = newest_dates)  %>% 
      ggsave(filename =  file.path(outputdir,
                                   paste0(cruisename, "-3d-plots"),
                                   paste0(cruisename, "-", tt, ".png")),
             width = 20, height = 7) 
  }
}

## Combine these things back into a csv-saveable table.
dates_list <- newest_dates %>% as.list()
recombined_list <- purrr::pmap(list(dates_list, newest_ylist, newest_biomass_list),
                               function(date_str, coords_matrix, qc_vector) {
  coords_tibble <- as_tibble(coords_matrix, .name_repair = "unique")
  colnames(coords_tibble) <- c("fsc_small_coord", "chl_small_coord", "pe_coord") # Rename columns
  coords_tibble %>%
    mutate(date = date_str, Qc = qc_vector)
})
hourly_gridded_reconstructed <- bind_rows(recombined_list) %>%
  select(date, fsc_small_coord, chl_small_coord, pe_coord, Qc)

## Each cruise is now saved to a file
filename = paste0(cruisename, "-gridded-after-cleaning.csv")
outfile = file.path(outputdir, filename)
write.csv(hourly_gridded_reconstructed, file = outfile, row.names = FALSE)
cat("wrote to", outfile, fill=TRUE)


## Example of how to load and reconstruct a |ylist| and |countslist| object.

# write.csv(hourly_gridded_reconstructed, file = outfile, row.names = FALSE)
# cat("wrote to", outfile, fill=TRUE)
```


### Cleaning Gradients2


```{r clean-g2}
#| fig-width: 12
#| fig-height: 12
#| output: false

## Load binned data
cruisename = "MGL1704"
datadir = "~/Dropbox/data/ocean/seaflow"
csv_filename = paste0(cruisename, "-gridded-before-cleaning-new.csv")
datlist <- csv_to_datlist(outputdir = "~/Dropbox/data/ocean/seaflow/old",
                          csv_filename = csv_filename)
ylist <- datlist$ylist
biomass_list <- datlist$biomass_list
dates <- datlist$datetimes

## Plotting before processing
gg_init = my_marginal_plotter(ylist, biomass_list, dates)

## One date we looked carefully at;
## dates[296+57], which is June 13th, 2017, 16:00

## Make a 3d particle plot to get a sense of how censoring has occurred
tt = 296 + 57
##tt = 50
# obj = plotly_particle(cruisename, tt, 1)
# print(obj)

## Remove top chl bin first (next up is removing bottom fsc_small)
res = my_edge_bin_remove(ylist, biomass_list, dimname = "chl_small", min_or_max = "max")
new_ylist = res$ylist
new_biomass_list = res$biomass_list
ylist_cens_after_chl = res$ylist_cens
biomass_list_cens_after_chl = res$biomass_list_cens
gg_after_chl = my_marginal_plotter(new_ylist, new_biomass_list, dates)
## cowplot::plot_grid(plotlist = list(gg_init,
##                                    gg_after_chl), ncol = 2)

## Remove top fsc too
res = my_edge_bin_remove(new_ylist, new_biomass_list, dimname = "fsc_small", min_or_max = "max")
newer_ylist = res$ylist
newer_biomass_list = res$biomass_list
ylist_cens_after_chl_and_fsc = res$ylist_cens
biomass_list_cens_after_chl_and_fsc = res$biomass_list_cens
gg_after_chl_and_fsc = my_marginal_plotter(newer_ylist, newer_biomass_list, dates)

## cowplot::plot_grid(plotlist = list(gg_init,
##                                    gg_after_chl,
##                                    gg_after_chl_and_fsc), ncol = 3)

combined_plot <-
cowplot::plot_grid(plotlist = list(gg_init %>% add_title("initial",.99),
                                   gg_after_chl %>% add_title("After CHL removal",.99),
                                   gg_after_chl_and_fsc %>% add_title("After CHL&FSC removal",.99)),ncol = 1) 
ggsave(filename =  file.path(outputdir,
                             paste0(cruisename, "-censor-removal.pdf")),
       plot = combined_plot,
       width = 10, height = 30)

## Make more 3d particle plots to get a sense of censoring
ttlist = c(18,155,226,97)
# plotlist = lapply(ttlist, function(tt){
#   ## tt = 18
#   ## tt = 155
#   ## tt = 226
#   ## tt = 97 ## Not an error, but interesting to see.
#   obj = plotly_particle(cruisename, tt, show_censored = TRUE)
#   obj
# })
# cowplot::plot_grid(plotlist = plotlist, ncol = 2)


## Make all 3d plots
if(FALSE){
  dir.create(file.path(outputdir, paste0(cruisename, "-3d-plots")))
  list.files(outputdir)
  TT = length(ylist)
  for(tt in 1:TT){
    print(tt)
    plot_3d_temp(ylist = newer_ylist,
                 biomass_list = newer_biomass_list,
                 tt = tt,
                 dates = dates)  %>% 
      ggsave(filename =  file.path(outputdir,
                                   paste0(cruisename, "-3d-plots"),
                                   paste0(cruisename, "-", tt, ".png")),
             width = 20, height = 7) 
  }
}

## Combine these things back into a csv-saveable table.
hourly_gridded_reconstructed = recombine_to_csv(newer_ylist, newer_biomass_list, dates)

## Each cruise's data is now saved to a file
filename = paste0(cruisename, "-gridded-after-cleaning.csv")
outfile = file.path(outputdir, filename)
write.csv(hourly_gridded_reconstructed, file = outfile, row.names = FALSE)
cat("wrote to", outfile, fill=TRUE)

## Also save the censored data
hourly_gridded_censored_reconstructed =
  recombine_censored_data(
      all_ylists = list(ylist_cens_after_chl, ylist_cens_after_chl_and_fsc),
      all_biomass_lists = list(biomass_list_cens_after_chl, biomass_list_cens_after_chl_and_fsc))
filename = paste0(cruisename, "-gridded-after-cleaning-censored.csv")
outfile = file.path(outputdir, filename)
write.csv(hourly_gridded_censored_reconstructed, file = outfile, row.names = FALSE)
cat("wrote to", outfile, fill=TRUE)
```



### Cleaning Gradients 1 

For G1, remove chl, then pe in order.

```{r clean-g1}
#| fig-width: 12
#| fig-height: 12
#| output: false

## Load binned data
cruisename = "SCOPE_16"
datadir = "~/Dropbox/data/ocean/seaflow"
csv_filename = paste0(cruisename, "-gridded-before-cleaning-new.csv")
datlist <- csv_to_datlist("~/Dropbox/data/ocean/seaflow/old",
                          csv_filename = csv_filename)
ylist <- datlist$ylist
biomass_list <- datlist$biomass_list
dates <- datlist$datetimes

## Plotting before processing
gg_init = my_marginal_plotter(ylist, biomass_list, dates)

## Make a 3d particle plot to get a sense of censoring
## tt = 100
## tt = 250
## tt = 310## Great example of censoring
tt = 97 ## Also great example of censoring
# obj = plotly_particle(cruisename, tt, 1)
# print(obj)

## Remove top chl (first; next up is removing bottom fsc_small)
res = my_edge_bin_remove(ylist, biomass_list, dimname = "chl_small", min_or_max = "max")
new_ylist = res$ylist
new_biomass_list = res$biomass_list
gg_after_chl = my_marginal_plotter(new_ylist, new_biomass_list, dates)


## Remove bottom PE
res = my_edge_bin_remove(new_ylist, new_biomass_list, dimname = "pe", min_or_max = "min")
newer_ylist = res$ylist
newer_biomass_list = res$biomass_list
gg_after_chl_and_pe = my_marginal_plotter(newer_ylist, newer_biomass_list, dates)

cowplot::plot_grid(plotlist = list(gg_init %>% add_title("initial",.99),
                                   gg_after_chl %>% add_title("After CHL removal",.99),
                                   gg_after_chl_and_pe %>% add_title("After CHL&PE removal",.99)),ncol = 1)  %>%
    ggsave(filename =  file.path(outputdir,
                                 paste0(cruisename, "-censor-removal.pdf")), width = 10, height = 30)


## Make all 3d plots
if(FALSE){
  dir.create(file.path(outputdir, paste0(cruisename, "-3d-plots")))
  list.files(outputdir)
  TT = length(ylist)
  for(tt in 1:TT){
    print(tt)
    plot_3d_temp(ylist = newer_ylist,
                 biomass_list = newer_biomass_list,
                 tt = tt,
                 dates = dates)  
      ggsave(filename =  file.path(outputdir,
                                   paste0(cruisename, "-3d-plots"),
                                   paste0(cruisename, "-", tt, ".png")),
             width = 20, height = 7) 
  }
}


## Combine these things back into a csv-saveable table.
dates_list <- dates %>% as.list()
recombined_list <- purrr::pmap(list(dates_list, newer_ylist, newer_biomass_list), function(date_str, coords_matrix, qc_vector) {
  coords_tibble <- as_tibble(coords_matrix, .name_repair = "unique")
  colnames(coords_tibble) <- c("fsc_small_coord", "chl_small_coord", "pe_coord") # Rename columns
  coords_tibble %>%
    mutate(date = date_str, Qc = qc_vector)
})
hourly_gridded_reconstructed <- bind_rows(recombined_list) %>%
  select(date, fsc_small_coord, chl_small_coord, pe_coord, Qc)

## Each cruise is now saved to a file
filename = paste0(cruisename, "-gridded-after-cleaning.csv")
outfile = file.path(outputdir, filename)
write.csv(hourly_gridded_reconstructed, file = outfile, row.names = FALSE)
cat("wrote to", outfile, fill=TRUE)
```


### Cleaning Gradients 4


```{r clean-g4}
#| fig-width: 12
#| fig-height: 12
#| output: false

## Load binned data
cruisename = "TN397_740"
datadir = "~/Dropbox/data/ocean/seaflow"
csv_filename = paste0(cruisename, "-gridded-before-cleaning-new.csv")
datlist <- csv_to_datlist("~/Dropbox/data/ocean/seaflow/old",
                          csv_filename = csv_filename)
ylist <- datlist$ylist
biomass_list <- datlist$biomass_list
dates <- datlist$datetimes

## Plotting before processing
gg_init = my_marginal_plotter(ylist, biomass_list, dates)

## It seems like we'll remove top row of  chl and fsc

## Make a 3d particle plot to get a sense of censoring
TT = length(ylist)
tt = 250
# obj = plotly_particle(cruisename, tt, 1)
# obj

## Remove top chl (first; next up is removing bottom fsc_small)
res = my_edge_bin_remove(ylist, biomass_list, dimname = "chl_small", min_or_max = "max")
new_ylist = res$ylist
new_biomass_list = res$biomass_list
gg_after_chl = my_marginal_plotter(new_ylist, new_biomass_list, dates)


## Remove top FSC 
res = my_edge_bin_remove(new_ylist, new_biomass_list, dimname = "fsc_small", min_or_max = "max")
newer_ylist = res$ylist
newer_biomass_list = res$biomass_list
gg_after_chl_and_fsc = my_marginal_plotter(newer_ylist, newer_biomass_list, dates)

combined_plot <- cowplot::plot_grid(plotlist = list(gg_init %>% add_title("initial",.99),
                                   gg_after_chl %>% add_title("After CHL removal",.99),
                                   gg_after_chl_and_fsc %>% add_title("After CHL&FSC removal",.99)),ncol = 1) 
ggsave(filename =  file.path(outputdir,
                             paste0(cruisename, "-censor-removal.pdf")),
       plot = combined_plot,
       width = 10, height = 30)

## Make all 3d plots
if(FALSE){
dir.create(file.path(outputdir, paste0(cruisename, "-3d-plots")))
list.files(outputdir)
TT = length(ylist)
for(tt in 1:TT){
  print(tt)
  plot_3d_temp(ylist = newer_ylist,
               biomass_list = newer_biomass_list,
               tt = tt,
               dates = dates)  
    ggsave(filename =  file.path(outputdir,
                                 paste0(cruisename, "-3d-plots"),
                                 paste0(cruisename, "-", tt, ".png")),
           width = 20, height = 7) 
}
}


## Combine these things back into a csv-saveable table.
dates_list <- dates %>% as.list()
recombined_list <- purrr::pmap(list(dates_list, newer_ylist, newer_biomass_list), function(date_str, coords_matrix, qc_vector) {
  coords_tibble <- as_tibble(coords_matrix, .name_repair = "unique")
  colnames(coords_tibble) <- c("fsc_small_coord", "chl_small_coord", "pe_coord") # Rename columns
  coords_tibble %>%
    mutate(date = date_str, Qc = qc_vector)
})
hourly_gridded_reconstructed <- bind_rows(recombined_list) %>%
  select(date, fsc_small_coord, chl_small_coord, pe_coord, Qc)

## Each cruise is now saved to a file
filename = paste0(cruisename, "-gridded-after-cleaning.csv")
outfile = file.path(outputdir, filename)
write.csv(hourly_gridded_reconstructed, file = outfile, row.names = FALSE)
cat("wrote to", outfile, fill=TRUE)
```


### Cleaning Gradients 5

```{r clean-g5}
#| fig-width: 12
#| fig-height: 12
#| output: false

## Load binned data
cruisename = "TN412_130" ## G5
datadir = "~/Dropbox/data/ocean/seaflow"
csv_filename = paste0(cruisename, "-gridded-before-cleaning-new.csv")
datlist <- csv_to_datlist("~/Dropbox/data/ocean/seaflow/old",
                          csv_filename = csv_filename)
ylist <- datlist$ylist
biomass_list <- datlist$biomass_list
dates <- datlist$datetimes

## Plotting before processing
gg_init = my_marginal_plotter(ylist, biomass_list, dates)

## It seems like we'll remove top row of  chl and fsc

## Make a 3d particle plot to get a sense of censoring
## tt = 100
TT = length(ylist)
tt = 100
tt = 36  ## Look at this next; are low PE particles also  high FSC (and high/low but not  mid CHL)?
# obj = plotly_particle(cruisename, tt, .1)
# print(obj)

## Remove top chl (first; next up is removing bottom fsc_small)
res = my_edge_bin_remove(ylist, biomass_list, dimname = "chl_small", min_or_max = "max")
new_ylist = res$ylist
new_biomass_list = res$biomass_list
gg_after_chl = my_marginal_plotter(new_ylist, new_biomass_list, dates)
cowplot::plot_grid(plotlist = list(gg_init,
                                   gg_after_chl), ncol = 2)


## Remove top FSC 
res = my_edge_bin_remove(new_ylist, new_biomass_list, dimname = "fsc_small", min_or_max = "max")
newer_ylist = res$ylist
newer_biomass_list = res$biomass_list
gg_after_chl_and_fsc = my_marginal_plotter(newer_ylist, newer_biomass_list, dates)
## cowplot::plot_grid(plotlist = list(gg_init,
##                                    gg_after_chl,
##                                    gg_after_chl_and_fsc), ncol = 3)
combined_plot = cowplot::plot_grid(plotlist = list(gg_init %>% add_title("initial",.99),
                                                   gg_after_chl %>% add_title("After CHL removal",.99),
                                                   gg_after_chl_and_fsc %>% add_title("After CHL&FSC removal",.99)),ncol = 1)
ggsave(filename = file.path(outputdir,
                            paste0(cruisename, "-censor-removal.pdf")),
       plot = combined_plot,
       width = 10, height = 30)

## Make all 3d plots
if(FALSE){
  dir.create(file.path(outputdir, paste0(cruisename, "-3d-plots")))
  list.files(outputdir)
  TT = length(ylist)
  for(tt in 1:TT){
    print(tt)
    plot_3d_temp(ylist = newer_ylist,
                 biomass_list = newer_biomass_list,
                 tt = tt,
                 dates = dates)  
    ggsave(filename =  file.path(outputdir,
                                 paste0(cruisename, "-3d-plots"),
                                 paste0(cruisename, "-", tt, ".png")),
           width = 20, height = 7) 
  }
}


## Combine these things back into a csv-saveable table.
dates_list <- dates %>% as.list()
recombined_list <- purrr::pmap(list(dates_list, newer_ylist, newer_biomass_list), function(date_str, coords_matrix, qc_vector) {
  coords_tibble <- as_tibble(coords_matrix, .name_repair = "unique")
  colnames(coords_tibble) <- c("fsc_small_coord", "chl_small_coord", "pe_coord") # Rename columns
  coords_tibble %>%
    mutate(date = date_str, Qc = qc_vector)
})
hourly_gridded_reconstructed <- bind_rows(recombined_list) %>%
  select(date, fsc_small_coord, chl_small_coord, pe_coord, Qc)

## Each cruise is now saved to a file
filename = paste0(cruisename, "-gridded-after-cleaning.csv")
outfile = file.path(outputdir, filename)
write.csv(hourly_gridded_reconstructed, file = outfile, row.names = FALSE)
cat("wrote to", outfile, fill=TRUE)
```
