---
title: "Data imputation and visualization for Environmental Factors"
author: "Sangwon Hyun"
date: "`r Sys.Date()`" 
execute:
  warning: false
  message: false
---

## 1. Environment Setup & Helper Functions

In this section, we load the necessary R packages and set up matrix visualizers.

```{r setup-packages}

## Core tidyverse packages
library(tidyverse)
library(ggplot2)
library(tidyr)
library(dplyr)
library(scales)

## Auxiliary packages
library(broom)
library(janitor)
library(softImpute)
library(kableExtra)

## Helper function for plotting matrix heatmaps
drawmat_precise_ggplot <- function(distmat, colours = c("blue", "red"),
                                   limits = NULL,
                                   xlab = "", ylab = "", title = ""){
  distmat <- as.matrix(distmat)
  longData <- reshape2::melt(distmat[nrow(distmat):1, ])

  p <- ggplot() +
    geom_raster(aes(x = Var2, y = Var1, fill = value), data = longData) +
    scale_fill_gradientn(colours = colours, guide = "colorbar", limits = limits) +
    labs(x = xlab, y = ylab, title = title) +
    theme_minimal() + 
    theme(
      axis.text.x = element_text(size = rel(1), angle = 90, vjust = 0.3),
      axis.text.y = element_text(size = rel(1)),
      plot.title = element_text(size = rel(1.5))
    )
  return(p)
}

#################
## Directories ##
#################


## contains files such as X-MGL1704.csv, from Dane's covariate_scraping scripts
## (Change to your Dropbox/data directory locally.)
datadir = "~/Dropbox/data/ocean/seaflow-covariates" 
##datadir = "~/Downloads" 

## Output directory for the imputed hourly covariates
## (Change to your Dropbox/data directory locally.)
## outputdir = "~/repos/process-data/output/seaflow-covariates" 
outputdir = "~/Dropbox/data/ocean/seaflow-covariates" 


#########################
## Helpers functions ####
#########################

cat_tab <- function(title, level = 3) {
  # The leading "\n\n" forces a clean break from whatever printed previously
  cat("\n\n", strrep("#", level), " ", title, "\n\n", sep = "")
}

make_html_table <- function(mytab, table_caption = NULL) {
  mytab |>
    kable(
      format = "html", 
      digits = 2, 
      align = "c",
      caption = table_caption
    ) |>
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE
    )
}
```

---

## 2. Data Loading & Standardization

Load the cruise dataset and scale numeric features.


```{r load-data}
#| code-fold: true


## I think these are "plain" names
## cruisenames_plain = c("SCOPE_16", ## G1
##                 "MGL1704", ## G2
##                 "KM1906", ## G3
##                 "TN397_740", ## G4
##                 "TN412_130") %>% ## G5
##   .[2] ## Only G2 for now.


cruisenames = c("KOK1606", ## G1
                "MGL1704", ## G2
                "KM1906", ## G3
                "TN397", ## G4
                "TN412")  ## G5

for(cruisename in cruisenames){

  ## Print the markdown header for the tab
  cat_tab(cruisename)

  ## Load data
  filename = paste0( cruisename, "-X.csv")
  X <- read.csv(file = file.path(datadir, filename))%>% as_tibble()
  
  ## Extract spatiotemporal identifiers and scale numerical variables
  time0 <- lubridate::as_datetime(X$time)
  lat0 <- X %>% pull(lat)
  lon0 <- X %>% pull(lon)
  varnames <- X %>% select(-c("time", "lat", "lon")) %>% colnames()
  
  ## 1. First take the means and standard deviations
  means = X %>% select(-time, -lat, -lon) %>%
    summarise(across(everything(), ~mean(.x, na.rm = TRUE))) 
  sds = X %>% select(-time, -lat, -lon) %>%
    summarise(across(everything(), ~sd(.x, na.rm = TRUE))) 
  
  ## Base R plot
  apply(X[-c(1,2,3)], 2, mean, na.rm=TRUE) %>% sort(decreasing=TRUE) %>%
    plot(main = "Means of all variables, sorted")
  cat("\n\n") 
  
  ## 2. Then, identify which variables have really big means
  thresh = 1E10
  somenames = means[,which(means>thresh & sds > thresh)] %>% names()
  if(length(somenames) >= 1){

    ## 3. Visualize 
    p <- X[,somenames] %>%
      add_column(time = time0) %>%
      pivot_longer(-time)  %>% 
      ggplot() +
      geom_point(aes(x=time, y = value, group = name), size = .5) +
      facet_wrap(~name, scale = 'free_y')   +
      ggtitle("The unusually high-value variables")
    
    print(p)
    cat("\n\n") 
    
    ## 4. Remove these variables
    X = X %>% select(-all_of(somenames))
  }
  
  ## Store the scaling information
  scaleobj <- X %>% select(-lat, -lon, -time) %>% scale() 
  mean_val <- attr(scaleobj, "scaled:center") %>% unname()
  sd_val <- attr(scaleobj, "scaled:scale") %>% unname()
  
  ## Double check that scaling is sound
  scaled_data = as.matrix(scaleobj)
  unscaled_data <- sweep(scaled_data, 2, sd_val, `*`)
  unscaled_data <- sweep(unscaled_data, 2, mean_val, `+`)
  orig_matrix <- as.matrix(X %>% select(-time, -lat, -lon))
  
  ## Base R plot
  max_diffs <- apply(abs(unscaled_data - orig_matrix), 2, max, na.rm = TRUE) %>% sort()
  plot(max_diffs, type = "p", xlab = "Variables (Sorted)",
       ylab = "Max Absolute Difference",
       main = "Reconstruction Error per Variable")
  abline(h = 0, col = "red", lty = 2)
  cat("\n\n") 
  
  ## Save the scaling information
  filename = paste0(cruisename, "-cmap-scaling-info.RDS")
  X_scaled = scaled_data
  saveRDS(list(X_before_scaling = X,
               X_after_scaling = X_scaled,
               scaleobj = scaleobj,
               X_means = mean_val,
               X_sds = sd_val),
          file = file.path(outputdir, filename))

  ## Plot the initial scaled variables across time: 
  # X_combined <- X_scaled %>% as_tibble() %>% mutate(time = time0,.before = 1)
  # scaled_plot <- X_combined %>% 
  #   pivot_longer(-time) %>% 
  #   ggplot() +
  #   geom_path(aes(x = time, y = value, group = name), alpha = 0.1, linewidth = 0.1) +
  #   theme_minimal() +
  #   ggtitle("Scaled variables over time")
  # print(scaled_plot)


  ## Analyze missingness patterns across all variables.
  ## > **Note:** Variables like salinity from GO-SHIP or GEOTRACES trajectories are naturally unavailable for non-targeted cruises ([GO-SHIP Paper](https://www.nature.com/articles/s41597-022-01212-w), [GEOTRACES Overview](https://www.geotraces.org/cruise-overview/)).


  ## Proportion of missing time points
  num_na <- apply(X_scaled, 2, function(mycol) sum(is.na(mycol))) 
  ##print(head(sort(num_na, decreasing = TRUE)))
  
  num_time_points <- nrow(X_scaled)
  prop_missing <- num_na / num_time_points
  
  ## Missingness distribution histogram
  hist(prop_missing, breaks = 30, main = "Proportion of missing time points")
  abline(v = 0.25, col = "red", lty = 2)
  
  missing_only_few <- sum(prop_missing < 0.25)
  legend("topleft", legend = paste0("Missing only\na few:\n", missing_only_few, "\nvariables"), box.lwd = 0)
  
  missing_a_lot <- sum(prop_missing > 0.25)
  legend("topright", legend = paste0("Missing \na lot:\n", missing_a_lot, "\nvariables"), box.lwd = 0)


  ## Visualize the feature matrix before imputation (variables with $<25\%$ missing values):
  X_scaled_copy0 <- X_scaled %>% as.matrix()
  dimnames(X_scaled_copy0) <- NULL
  
  # heatmap <- X_scaled_copy0 %>% 
  #   .[, prop_missing < 0.25] %>% 
  #   drawmat_precise_ggplot() + 
  #   ggtitle("Matrix heatmap before imputation (<25% missing)")
  # print(heatmap)

  ## Matrix Completion via Low-Rank SVD (`softImpute`); we'll
  ## fill missing values using nuclear-norm regularized singular value decomposition (`softImpute`).
  X_scaled_copy <- X_scaled_copy0[, prop_missing < 0.25] 

  if(sum(is.na(X_scaled_copy)) >= 1){

    ## Fit nuclear-norm regularized SVD model
    fit <- softImpute(X_scaled_copy, rank.max = 100, lambda = 10) ## I haven't carefully chosen lambda (or rank.max); we should do this..
    
    ## Complete matrix (fill NAs using low-rank fit)
    X_hat <- complete(X_scaled_copy, fit, unscale = FALSE)

    ## Visualize the result
    # heatmap2 <- drawmat_precise_ggplot(X_hat) +
    #   ggtitle("Matrix heatmap after softImpute low-rank completion")
    # print(heatmap2)

    ## Compare imputed entries against original observed series:
    X_smaller <- X_scaled_copy
    observed_vec <- as.vector(X_smaller)
    p <- ncol(X_smaller)
    n <- nrow(X_smaller)
    
    df <- tibble(
      Time = rep(1:n, times = p),
      Series = rep(1:p, each = n),
      Value = as.vector(X_hat),
      Imputed_or_not = is.na(observed_vec)  
    )
    
    df_plot <- df %>%
      mutate(
        Series_Label = paste("Series", Series),
        Status = ifelse(Imputed_or_not, "Imputed", "Observed")
      )
    
    imputed_plot <- df_plot %>%
      group_by(Series) %>%
      filter(any(Imputed_or_not)) %>%
      ggplot(aes(x = Time, y = Value)) +
      geom_line(color = "gray50", linewidth = 0.5, alpha = 0.7) +
      geom_point(
        data = . %>% filter(Imputed_or_not),
        aes(color = Status),
        size = 0.3
      ) +
      facet_wrap(~ Series_Label, scales = "free_y") +
      scale_color_manual(
        values = c("Imputed" = "firebrick", "Observed" = "gray30")
      ) +
      labs(
        title = "Time Series Imputation via softImpute",
        ##| fig-cap: "Comparison of observed vs. softImpute filled values across series"
        x = "Time Index",
        y = "Value",
        color = "Data Point"
      ) +
      theme_minimal() +
      theme(
        legend.position = "bottom",
        strip.text = element_text(face = "bold")
      )

    print(imputed_plot)

  } else {

    X_hat = X_scaled_copy0[, prop_missing < 0.25]

  }

  ## Sanity check: confirm non-imputed entries were untouched:
  (X_scaled_copy0[, prop_missing < 0.25] - X_hat) %>% 
    as.numeric() %>% 
    summary() %>%
    as.list() %>%
    as_tibble(check.names = FALSE)  %>%
    make_html_table("Summary of the reconstruction errors") %>% print()


  ## Format the imputed output, re-attach spatial coordinates, and aggregate to hourly resolution.
  varnames = colnames(X_scaled)
  remaining_names = varnames[prop_missing < 0.25]
  colnames(X_hat) <- remaining_names
  
  X_imputed <- X_hat %>% 
    as_tibble() %>% 
    add_column(time = time0, lat = lat0, lon = lon0, .before = 1)
  
  X_imputed_hourly <- X_imputed %>% 
    group_by(time = lubridate::floor_date(time, "hour")) %>%
    summarize(across(where(is.numeric), mean), .groups = "drop")

  # hourly_plot <- X_imputed_hourly %>% 
  #   select(-lat, -lon) %>%
  #   pivot_longer(-time) %>%
  #   ggplot() +
  #   geom_line(aes(x = time, y = value, group = name), col = "blue", alpha = 0.2, linewidth = 0.1) +
  #   scale_x_datetime(
  #       breaks = "1 day",
  #       labels = label_date("%b %d\n00:00") 
  #   ) +
  #   theme_minimal() +
  #   theme(panel.grid.minor.x = element_blank()) +
  #   ggtitle("Hourly aggregated variables over time")
  # print(hourly_plot)

  # hourly_plot_over_lat <- X_imputed_hourly %>% 
  #   pivot_longer(-c(time, lat, lon)) %>% 
  #   ggplot(aes(x = lat, y = value, group = name)) +
  #   geom_path(col = "blue", linewidth = 0.1, alpha = 0.2) +
  #   theme_minimal() +
  #   ggtitle("Hourly aggregated variables plotted across latitude")
  # print(hourly_plot_over_lat)

  ## Save the imputed covariates in a single file;
  ## Saves to MGL1704-X-from-cmap-imputed-and-scaled.csv

  filename = paste0(cruisename, "-X-from-cmap-imputed-and-scaled.csv")
  write.csv(X_imputed_hourly,
            file = file.path(outputdir, filename),
            row.names = FALSE)
}

## Close the tabset from INSIDE the chunk
cat("\n\n:::\n\n")
```

::: {.panel-tabset}

```{r load-data-actual}
#| ref.label: "load-data"
#| echo: false
#| results: asis
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 5
```


---


```{r}
# Stop rendering the document right here
knitr::knit_exit()
```




---

## 7. Principal Component Analysis (PCA)

Run PCA on observation time points across all imputed variables.

```{r pca-analysis}

pca_res <- X_imputed_hourly %>% 
  select(-time, -lat, -lon) %>% 
  prcomp(scale. = TRUE) %>%
  broom::augment(X_imputed_hourly %>% select(time, lat, lon)) %>% 
  janitor::clean_names() %>% 
  rename_with(~ gsub("^fitted_", "", .x), starts_with("fitted_")) %>%
  select(time, lon, lat, matches("^pc[1-6]$"))
```

```{r fig-pca-components}
#| fig-width: 10
#| fig-height: 6
#| fig-cap: "First six principal components across time"

pca_res %>% 
  select(-lat, -lon) %>% 
  pivot_longer(-time) %>%
  ggplot() +
  geom_line(aes(x = time, y = value, group = name)) +
  facet_wrap(~ name) +
  theme_minimal()
```

---

## 7. Next Directions

1. Integrate onboard lab variables (PAR, salinity, temperature).
2. Perform unsupervised variable space embeddings (PCA/UMAP on transposed feature matrix).
