# Ocean Environmental Covariates for Seaflow cruises

## Loading and visualizing data

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(tidyverse) 
library(cmap4r)
library(cruisedat)

## Output directory
here::i_am("01-process-seaflow.Rmd") ## setwd("~/repos/process-data")
outputdir = "output/seaflow-covariates" ##outputdir = "~/Dropbox/data/ocean/seaflow/2025-09-15"
dir.create(outputdir, showWarnings = FALSE)

## Helpers
## source("02-helpers.R")

## Other setup
## colocalized_csvfile = file.path(outputdir, paste0(cruisename, "-from-CMAP.csv"))
```

We'll procure the environmental covariates that go along with seaflow cruises,
from Simons CMAP.


What are all the tables and variables we need?

```{r, eval = FALSE}
## Get authorization
cmap4r::set_authorization(reset = FALSE, cmap_key = "ec7b8400-e5ae-11eb-88c6-c79daf4f0e4a")

## Get the cadtalog
catalog = cmap4r::get_catalog()

## Do colocalization 
## cmap4r::get_cruises()
targetVars = catalog %>% filter(Table_Name == "tblPisces_NRT") %>% pull(Variable)
dat = cmap4r::along_track("MGL1704", targetTables = "tblPisces_NRT",
                    targetVars = "Fe", temporalTolerance = 3.5, latTolerance = 5,
                    lonTolerance = 5, depthTolerance = 10, depth1 = 0, depth2 = 10) 
dat %>% ggplot() + geom_point(aes(x=time, y= Fe))

## This will further be averaged into unique hours.
dat %>% mutate(hourly_time = lubridate::floor_date(time, unit = "hours")) %>%
  group_by(hourly_time) %>%
  summarize(Fe = mean(Fe, na.rm=TRUE)) %>% 
ggplot() + geom_point(aes(x=hourly_time, y= Fe))

## Then, the coverage is assessed.
dat %>% ggplot() + geom_point(aes(y=lat, x=lon, col = Fe))

## repodir = "~/repos/cruisedat"
## outputdir = "."
## dir.create(file.path(".", "export"), showWarnings = FALSE)
```

(It may not be practical to simply pull all data from all tables, as I used to
back in 2019.)

I realize there are at least three types of covariates.

- **Collected directly on the ship.** Each cruise's webpage shows a list of
  these. One probably needs to go over these manually, and selectively download
  them.
  + Also, there are measurements that are taken *sparsely* along a cruise track,
    sometimes at depth. These are only useful if they can be reasonably
    interpolated from the original resolution to the hourly level.

- **In situ drifter/float**
  + Global Drifter Program: (1979-02-15 – 2019-08-07): `tblGlobalDrifterProgram`
  + Argo Float Core Profiles (1997-07-28 – 2024-06-09): `tblArgoCore_REP_Jun2024`
    + I'm guessing this is the latest updated version. Others include:
    `tblArgoCore_REP_Feb2024`, `tblArgoCore_REP_Nov2023`, `tblArgoCore_REP`
  + Argo Float Biogeochemical Profiles (2002-09-08 – 2024-06-08)  `tblArgoBGC_REP_Jun2024`
    + I'm guessing this is the latest updated version. Others include:
    `tblArgoBGC_REP`, `tblArgoBGC_REP_Feb2024`, `tblArgoBGC_REP_Nov2023`,
    `tblArgoBGC_REP_Jun2023`
  + GEOTRACES Intermediate Data Product 2021 v2 - Seawater (2006-07-13 – 2018-11-22): `tblGeotraces_Seawater_IDP2021v2`
  + UCSD Argo Mixed Layer Depth Climatology Model 2022: `tblArgo_MLD_2022_Climatology` 
    + Seems super relevant
    + Also similar to `tblArgo_MLD_Climatology`

- **Models (climatology)**
  + Mercator-Pisces Biogeochemistry and Weekly Forecast  (2011-12-31 to 2019-04-27): `tblPisces_NRT` 
  + GOBAI-O2 Climatology v2.2 (Gridded Ocean Biogeochemistry from Artificial Intelligence - Oxygen)
  + Mercator-Pisces Biogeochemistry Daily Forecast Nutrients (2022-01-01 to 2025-07-18): `tblPisces_Forecast_Nutrients`
  + Darwin Biogeochemistry 3 Day Averaged Model (1994-01-03 to 2015-12-30): `tblDarwin_Nutrient`
  + World Ocean Atlas 2013 v2 Climatology: `tblWOA_Climatology`
  + MODIS Chlorophyll Concentration (2002-07-04 to 2022-06-10): `tblModis_CHL`
  + MODIS Photosynthetically Available Radiation Daily 9km (2010-01-01 to 2022-03-07): `tblModis_PAR`

- **Other less promising models**
  + Mercator-Pisces Biogeochemistry Daily Forecast Car: (2022-01-01 – 2025-07-18): `tblPisces_Forecast_Car`
  + Mercator-Pisces Biogeochemistry Daily Forecast Pft: (2022-01-01 – 2025-07-18): `tblPisces_Forecast_Pft`
  + GOBAI-O2 Climatology v2.2 Gridded Ocean Biogeochemistry from Artificial
    Intelligence - Oxygen (UNknown time range): `tblGOBAI_O2_v2_2_Climatology`
  + Mercator-Pisces Biogeochemistry Daily Forecast Co2 (2022-01-01 to 2025-07-18): `tblPisces_Forecast_Co2`

So, a strategy is to go through these steps:
- Manually co-localize the above tables and all sensible variables therein. 
- Examine the on-cruise measurements (from, e.g., https://simonscmap.com/catalog/cruises/MGL1704 for Gradients 2), by 
  1. Excluding any gated products (which will have lots of observations, thousands or tens of thousands), 
  2. Exclude any things that are related to other non-Seaflow cytometers (IFCB,
     or imaging flow cytobot, which measures large particles above 3microns,
     which is larger than the small picophytoplankton we want to study),
  3. Exclude any datasets that have too few surface observations (fewer than 30
     rows altogether, or in the "DEPTH" visualization tab, fewer than 30-ish
     observations on the surface.)
  4. Exclude GRUMP altogether.

Next, for every model/climatology or drift/floater dataset, we need to use a for-loop that goes over a time radius of {1 hr, 1 day, 3.5 days, 14 days}, and a latitude+longitude box of side lengths {0.1 degrees, 1 degrees, 2 degrees, 5 degrees}. 

(I have some code lying around (in github repository `sangwon-hyun/many-cruises`) that can do this.)

For now, this code is borrowed from `02-covariates`.

```{r, eval = FALSE}
base = "00-covariates"
source(here::here("data", base, "colocalize-params.R"))

## Temporarily placing here (redundant)
ids_final = readRDS(file = here::here("data", base, "ids_final.RDS"))
vars = read.csv(file = file.path("/home/sangwonh/repos/cruisedat", "CMAP_vars_all_cruises.csv")) %>% as_tibble()
vars = vars %>% mutate(Table_Name = str_replace_all(Table_Name,
                                             pattern = "tblWind_NRT",
                                             replacement = "tblWind_NRT_deprecated")) ##%>% kable() %>% print()
vars %>% kable() %>% print()
nvar = nrow(vars)
## End of temporary

datlist = mclapply(ids_final, function(id){

  ## Safely restarting cruises that don't exist
  if(file.exists(file = file.path(outputdir, "cmap-data", paste0(id, "-adaptive.RDS")))) return(NULL)

  ## Colocalize all variables
  reslist = lapply(1:nvar, function(ivar){
    tryCatch({

      varname = vars %>% pull(Variable) %>% .[ivar]
      tabname = vars %>% pull(Table_Name) %>% .[ivar]

      if(tabname %in% c("tblWOA_Climatology", "tblPisces_NRT", "tblCHL_REP", "tblModis_PAR")){
        param = list_of_params[[tabname]]
      } else {
        param = list_of_params[["other"]]
      }
      for(isize in 1:5){

        ## Increase the radius
        if(isize > 1) param$latTolerance = param$latTolerance + 0.1 * 4
        if(isize > 1) param$lonTolerance = param$lonTolerance + 0.1 * 4

        ## Perform the colocalization
        dt = along_track(id,
                         targetTables = tabname,
                         targetVars = varname,
                         temporalTolerance = param$temporalTolerance,
                         latTolerance = param$latTolerance,
                         lonTolerance = param$lonTolerance,
                         depthTolerance = param$depthTolerance,
                         depth1 = param$depth1,
                         depth2 = param$depth2) %>% as_tibble() 
        dt = dt %>% dplyr::select(!contains("_std"))

        ## Check the missing proportion
        dt_missing_prop = dt %>% summarize(naprop = mean(is.na(!!sym(varname)))) %>% unlist()
        ## print(dt_missing_prop)
        if(dt_missing_prop == 0) break
      }
      return(dt)
    }, error = function(e){
      return(NULL)
    })
  })
  lens = reslist %>% purrr::map(. %>% length()) %>% unlist()
  if(any(lens == 0)){ reslist = reslist[-which(lens == 0)] }
  if(length(reslist) > 0){
    fullres = reslist %>% purrr::reduce(full_join, by = c("time", "lon", "lat"))
    fullres = fullres %>% dplyr::select(!contains("_std"))
    saveRDS(fullres, file = file.path(outputdir, "cmap-data", paste0(id, "-adaptive.RDS")))
    ## saveRDS(fullres, file = file.path(outputdir, "cmap-data", paste0(id, "-adaptive.RDS")))
    return(NULL)
  }
  return(NULL)
}, mc.cores = length(ids_final), mc.preschedule = FALSE)
```



This is what I *used* to do; the CMAP_vars are variables that were used for the flowmix paper.

```{r, eval = FALSE}
## This is the *superset* of all the relevant data tables and variables
dat = read.csv(file = file.path(outputdir, "CMAP_vars.csv")) ## This needs to be updated.

## Get the duplicated
dat = unique(dat[,c("Variable", "Table_Name")])
dat = apply(dat, 2, as.character)

## Make some manual changes
dat[15, "Table_Name"] = "tblCHL_REP"
dat = dat[-which(dat[,"Variable"] %in% c("disp_bw_adt", "ftle_bw_adt")),]
## ADT and SLA are different, per:
## https://www.aviso.altimetry.fr/en/data/data-access/las-live-access-server/lively-data/2004/june-15-2004-sla-vs-adt.html
## They seem to have removed adt from the dataset; it doesn't exist.

## Some basic checks
assert_that(all(check_with_catalog(dat[,"Variable"],"varName",
                                   apiKey = apiKey)))
assert_that(all(check_with_catalog(dat[,"Table_Name"],"tableName",
                                   apiKey = apiKey)))

## Not sure why salinity is not in here, but include it.
dat = rbind(c("sss", "tblSSS_NRT"), dat)
varNames = dat[,"Variable"]
tableNames = dat[,"Table_Name"]
```

