---
title: "Adding on-board covariates"
author: "Sangwon Hyun"
date: "`r Sys.Date()`" 
execute:
  warning: false
  message: false
---

## 1. Environment Setup & Helper Functions

In this section, we load the necessary R packages and set up matrix visualizers.

```{r} 
#| label: setup-packages

## Core tidyverse packages
library(tidyverse)
library(ggplot2)
library(tidyr)
library(dplyr)
library(scales)

## Auxiliary packages
library(broom)
library(janitor)
library(softImpute)
library(kableExtra)

#################
## Directories ##
#################

## contains files such as X-MGL1704.csv, from Dane's script
## (Change to your Dropbox/data directory locally.)
datadir = "~/Dropbox/data/ocean/seaflow-covariates" 
##datadir = "~/Downloads" 

## Output directory for the imputed hourly covariates
## (Change to your Dropbox/data directory locally.)
## outputdir = "~/repos/process-data/output/seaflow-covariates" 
outputdir = "~/Dropbox/data/ocean/seaflow-covariates" 

```

I'll introduce an interpolation function for sunlight, since the values are
missing in several of the gradients cruises.

Specifically, I'll fill in from *the same time* on the surrounding days (e.g.,
±24h, ±48h).

(I'll also linearly fill in the other covariates, for completeness.)



```{r impute-helpers}
##' Fill in missing sunlight (PAR) and covariates using multi-day offsets.
##'
##' First ensures a continuous hourly time grid, then imputes missing
##' Photosynthetically Active Radiation (\code{par}) values by averaging
##' measurements at exact multi-day offsets (with optional hourly slack), and
##' linearly interpolates other numeric columns specifically at the newly
##' imputed PAR timestamps.
##'
##' @param geodat A data frame or tibble containing a datetime column named \code{time}, 
##'   a numeric column named \code{par}, and any other relevant numeric covariates.
##' @param days Numeric. The number of days before and after to target for strict 
##'   matching (e.g., \code{days = 1} checks $\pm 24$h, \code{days = 2} checks $\pm 24$h and $\pm 48$h). 
##'   Default is `1`.
##' @param slack Numeric. The number of hours of slack (buffer window) to add and subtract 
##'   around each day offset. Default is `0` (strict exact offsets only).
##'
##' @return A tibble with a complete hourly sequence of timestamps from the minimum to 
##'   maximum time range, containing the updated \code{par_imputed} column and linearly 
##'   interpolated values for other numeric variables at the imputed PAR rows.
##'
##' @import dplyr
##' @importFrom zoo na.approx
##' @export
fill_in_par <- function(geodat, days = 1, slack = 0) {
  
  ## 0. Ensure a continuous hourly grid from the earliest to latest time
  geodat <- tibble(
    time = seq(min(geodat$time, na.rm = TRUE), 
               max(geodat$time, na.rm = TRUE), 
               by = "1 hour")
  ) %>% 
    left_join(geodat, by = "time")
  
  n <- nrow(geodat)
  
  ## 1. Generate strict hour offsets (e.g., ±24h, ±48h)
  day_offsets <- c(seq(-days, -1), seq(1, days))
  strict_hour_offsets <- day_offsets * 24
  
  ## 2. Generate slack hour offsets
  slack_hours <- seq(-slack, slack, by = 1)
  all_hour_offsets <- unlist(lapply(strict_hour_offsets, function(d_hr) {
    d_hr + slack_hours
  })) %>% unique()
  
  ## Helper to pull values at specific hour shifts for every row simultaneously
  get_offset_matrix <- function(hour_offsets) {
    sapply(hour_offsets, function(h) {
      row_indices <- seq_len(n) + h
      valid <- row_indices >= 1 & row_indices <= n
      res <- rep(NA_real_, n)
      res[valid] <- geodat$par[row_indices[valid]]
      res
    })
  }
  
  strict_mat <- get_offset_matrix(strict_hour_offsets)
  slack_mat  <- get_offset_matrix(all_hour_offsets)
  
  ## 3. Compute row means
  impute_strict <- rowMeans(strict_mat, na.rm = TRUE)
  impute_strict[is.nan(impute_strict)] <- NA_real_
  
  impute_slack <- rowMeans(slack_mat, na.rm = TRUE)
  impute_slack[is.nan(impute_slack)] <- NA_real_
  
  impute_val <- coalesce(impute_strict, impute_slack)
  
  ## 4. Extract the exact contributing values used for each row as a clean list-column
  ## (Pulls the row from strict_mat if available, otherwise uses slack_mat)
  contributing_list <- lapply(seq_len(n), function(i) {
    if (!is.na(impute_strict[i])) {
      vals <- strict_mat[i, ]
    } else {
      vals <- slack_mat[i, ]
    }
    vals[!is.na(vals)] ## Keep only valid, non-NA contributing values
  })
  
  ## 5. Assemble final data frame
  geodat <-
    geodat %>% 
    mutate(
      contributing_values = contributing_list,
      par_imputed = coalesce(par, impute_val)
    ) %>%
    select(-contributing_values)


  ## 6. Fills in the rest of the values using linear interpolation.
  geodat = geodat %>%
    mutate(
    is_par_imputed = is.na(par) & !is.na(par_imputed),
    across(
      where(is.numeric) & !matches("time|par|par_imputed|is_par_imputed"),
      ~ {
        interpolated <- zoo::na.approx(.x, na.rm = FALSE)
        ifelse(is_par_imputed, interpolated, .x)
      }
    )
  ) %>% 
  select(-is_par_imputed)
  
  return(geodat)
}


##' Visualize Imputed vs. Original PAR
##'
##' Generates a ggplot displaying the original Photosynthetically Active 
##' Radiation (\code{par}) values alongside the newly imputed version 
##' (\code{par_imputed}) over time.
##'
##' @param geodat A data frame or tibble containing \code{time}, \code{par}, 
##'   and \code{par_imputed} columns (typically the output of \code{fill_in_par()}).
##' @param cruisename Character string. Optional name of the cruise to display 
##'   in the plot title. Default is an empty string.
##'
##' @return A \code{ggplot} object showing the overlay of original and imputed PAR.
##'
##' @import ggplot2
##' @export
make_par_plot <- function(geodat, cruisename = ""){
  par_plot <- geodat %>% 
    ggplot() +
    geom_line(aes(x=time, y=par_imputed), col = 'red', size = rel(2)) +
    geom_line(aes(x=time, y = par), col = 'black', size = rel(2)) +
      ggtitle(paste0("Imputing PAR for ", cruisename))  +
    scale_x_datetime(
      date_breaks = "1 day", 
      date_labels = "%b %d" # Format as "Month Day" (e.g., Sep 04), change to "%Y-%m-%d" if you want the year
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
  return(par_plot)
}

## Color scale for line plots
custom_colors <- c(
    "lat"          = "black",
    "lon"          = "grey50",
    "temp"         = "#E64B35", # Warm red
    "par"          = "#F39C12", # Gold / yellow
    "salinity"     = "#2980B9", # Blue
    "conductivity" = "#8E44AD"  # Purple
)
onboard_vars = names(custom_colors) 
```


---

## Loading and cleaning covariates collected aboard the cruises

Load the handful of environmental covariates from on-board the cruise.

```{r load-onboard}
#| fig-width: 20
#| fig-height: 12


raw_data_dir <- "~/Dropbox/data/ocean/raw-seaflow/particle-data-2025-11-21"

list.files(raw_data_dir)
cruises = c("SCOPE_16", ## G1
            "MGL1704", ## G2
            "KM1906", ## G3
            "TN397_740", ## G4
            "TN412_130") ## G5

datlist = lapply(cruises, function(cruisename){

  ## Load sunlight (par), temperature, conductivity, salinity
  geodat =
    file.path(raw_data_dir, cruisename, "dash", paste0(cruisename, "-geo.tsdata"))   %>%
    read.csv(sep = "\t", skip = 6)  %>% as_tibble() %>%
    mutate(time = lubridate::as_datetime(time))
  
  geodat =
    geodat %>%
    mutate(hour_group = floor_date(time, "1 hour")) %>%
    group_by(hour_group) %>%
    summarise(across(where(is.numeric), \(x) mean(x, na.rm=TRUE))) %>%
    rename(time = hour_group)
  
  ## Manual adjustments
  if(cruisename == "MGL1704"){


    geodat = geodat %>%
      mutate(
          ## 1. Turn the bad value into NA
          salinity = if_else(time %in% 
                             c(lubridate::as_datetime("2017-06-05 01:00:00"),
                               lubridate::as_datetime("2017-06-05 02:00:00"),
                               lubridate::as_datetime("2017-06-05 03:00:00"),
                               lubridate::as_datetime("2017-06-05 04:00:00"),
                               lubridate::as_datetime("2017-05-30 00:00:00")
                               ), NA_real_, salinity),
    # 2. Automatically linearly interpolating missing values based on the time vector
    salinity = zoo::na.approx(salinity, x = time, na.rm = FALSE)
    )

    geodat = geodat %>%
      mutate(
          ## 1. Turn the bad value into NA
          conductivity = if_else(time %in% 
                             c(lubridate::as_datetime("2017-06-05 01:00:00"),
                               lubridate::as_datetime("2017-06-05 02:00:00"),
                               lubridate::as_datetime("2017-06-05 03:00:00"),
                               lubridate::as_datetime("2017-06-05 04:00:00")), NA_real_, conductivity),
    # 2. Automatically interpolate missing values based on the time vector
    conductivity = zoo::na.approx(conductivity, x = time, na.rm = FALSE)
    )


    ## Impute PAR (using exactly +/- 24 , 48, ... hours)
    geodat =
      geodat |> fill_in_par(days = 5, slack = 2) 
    geodat %>% make_par_plot(cruisename) %>% print()
    geodat = geodat %>% select(-par) %>% rename(par = par_imputed)

  }

  if(cruisename == "SCOPE_16"){
    ## Clean one single point using the average of the +-1 hours
    t_target <- lubridate::as_datetime("2016-05-02 05:00:00")
    t_prev   <- lubridate::as_datetime("2016-05-02 04:00:00")
    t_next   <- lubridate::as_datetime("2016-05-02 06:00:00")
    geodat$salinity[geodat$time == t_target] <- 
      (geodat$salinity[geodat$time == t_prev] + geodat$salinity[geodat$time == t_next]) / 2

    ## Impute PAR (using exactly +/- 24, 48, and 72 hours)
    geodat = geodat |> fill_in_par(days = 3)
    geodat %>% make_par_plot(cruisename) %>% print()
    geodat = geodat %>% select(-par) %>% rename(par = par_imputed)
  }

  ## Manual adjustments
  if(cruisename == "KM1906"){
    geodat = geodat %>% filter(time > lubridate::as_datetime("2019-04-10 12:00:00"))  

    ## Clean one single point using the average of the +-1 hours
    t_target <- lubridate::as_datetime("2019-04-15 03:00:00")
    t_prev   <- lubridate::as_datetime("2019-04-15 02:00:00")
    t_next   <- lubridate::as_datetime("2019-04-15 04:00:00")
    geodat$salinity[geodat$time == t_target] <- 
      (geodat$salinity[geodat$time == t_prev] + geodat$salinity[geodat$time == t_next]) / 2

    ## Impute PAR (using exactly +/- 24 hours)
    geodat = geodat |> fill_in_par(days = 2)
    geodat %>% make_par_plot(cruisename) %>% print()
    geodat = geodat %>% select(-par) %>% rename(par = par_imputed)
  }


  if(cruisename %in% c("TN412_130", "TN397_740")){

    ## Impute PAR (using exactly +/- 24, 48, and 72 hours)
    geodat = geodat |> fill_in_par(days = 3)
    geodat %>% make_par_plot(cruisename) %>% print()
    geodat = geodat %>% select(-par) %>% rename(par = par_imputed)

  }


  return(geodat)
})
names(datlist) = cruises


## Make plots
plotlist <- lapply(cruises, function(cruisename){
  geodat = datlist[[cruisename]]
  geodat %>%
    pivot_longer(-time) %>%
    mutate(name = factor(name, levels = c("lat", "lon", "temp", "par", "salinity", "conductivity"))) %>% 
    ggplot(aes(x = time, y = value, group = name, color = name)) +
    geom_line(linewidth = 0.8) +
    facet_wrap(~name, scales = "free_y", ncol = 6) +
    ggtitle(cruisename) +
    scale_color_manual(
      values = custom_colors,
      limits = c("lat", "lon", "temp", "par", "salinity", "conductivity"),
      drop = FALSE
    ) +
    theme_minimal()
})
cowplot::plot_grid(plotlist = plotlist, nrow =5)


## Combine them
plotlist <- lapply(names(datlist), function(nm) {
  dat <- datlist[[nm]]
  drawmat_precise_ggplot(dat %>% select(-c("time")) %>% as.matrix()%>% scale()) +
    ggtitle(nm)
})
cowplot::plot_grid(plotlist = plotlist, ncol = 5)

combined_dat = datlist %>% bind_rows(.id = "cruisename") %>% rename(temp_onboard = temp,
                                                                    salinity_onboard = salinity,
                                                                    par_onboard = par,
                                                                    conductivity_onboard = conductivity)
```


---

## Scale and save the on-board data


```{r scale-and-save-on-board}
#| fig-width: 12
#| fig-height: 7

for(cruisename in cruises){

  ## Take unscaled data
  X = datlist[[cruisename]]


  ## Remove some timepoints 
  if(cruisename == "MGL1704"){
    X = X %>% filter(time > lubridate::as_datetime("2017-05-28 15:00:00"))
  }

  ## Save time/lat/lon.
  time0 <- lubridate::as_datetime(X$time)
  lat0 <- X %>% pull(lat)
  lon0 <- X %>% pull(lon)

  ## Store the scaling information
  scaleobj <- X %>% select(-lat, -lon, -time) %>% scale() 
  mean_val <- attr(scaleobj, "scaled:center") %>% unname()
  sd_val <- attr(scaleobj, "scaled:scale") %>% unname()
  
  ## Double check that scaling is sound
  scaled_data = as.matrix(scaleobj)
  unscaled_data <- sweep(scaled_data, 2, sd_val, `*`)
  unscaled_data <- sweep(unscaled_data, 2, mean_val, `+`)
  orig_matrix <- as.matrix(X %>% select(-time, -lat, -lon))
  max_diffs <- apply(abs(unscaled_data - orig_matrix), 2, max, na.rm = TRUE) %>% sort()
  stopifnot(all(max_diffs < 1E-8))
  
  ## Save the scaling information
  filename = paste0(cruisename, "-on-board-scaling-info.RDS")
  X_scaled = scaled_data
  saveRDS(list(X_before_scaling = X,
               X_after_scaling = X_scaled,
               scaleobj = scaleobj,
               X_means = mean_val,
               X_sds = sd_val),
          file = file.path(outputdir, filename))

  ## Save the scaled data
  X_combined <- X_scaled %>% as_tibble() %>% mutate(time = time0, lat = lat0, lon = lon0, .before = 1)
  filename = paste0(cruisename, "-X-on-board.csv")
  write.csv(X_combined, file = file.path(outputdir, filename), row.names = FALSE)

  ## Make a plot, just because
  scaled_plot <-
    X_combined  %>% ##%>% select(-lat) %>% select(-lon) %>% 
    mutate(across(where(is.numeric) & -time, ~ as.numeric(scale(.x)))) %>% 
    pivot_longer(-time) %>% 
    ggplot() +
    geom_path(aes(x = time, y = value, group = name, col = name), alpha = 1, linewidth = 1) +
    theme_minimal() +
    ggtitle(paste0(cruisename, ": Scaled variables over time")) +
    scale_color_manual(
      values = custom_colors,
      limits = c("lat", "lon", "temp", "par", "salinity", "conductivity"),
      drop = FALSE
    )  +
    scale_x_datetime(
      date_breaks = "1 day", 
      date_labels = "%b %d" # Format as "Month Day" (e.g., Sep 04), change to "%Y-%m-%d" if you want the year
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
  print(scaled_plot)
}
```



---

## Combine with CMAP-downloaded covariates

```{r combine-cmap-and-onboard}

cruises_cmap = c(SCOPE_16 = "KOK1606", ## G1
                 MGL1704 = "MGL1704", ## G2
                 KM1906 = "KM1906", ## G3
                 TN397_740 = "TN397", ## G4
                 TN412_130 = "TN412") ## G5

cruises = c("SCOPE_16", ## G1
            "MGL1704", ## G2
            "KM1906", ## G3
            "TN397_740", ## G4
            "TN412_130") ## G5

for(cruisename in cruises){ 

  ## Load cmap data
  cruisename_cmap = cruises_cmap[cruisename]
  filename = paste0(cruisename_cmap, "-X-from-cmap-imputed-and-scaled.csv")
  X_cmap = read.csv(file = file.path(outputdir, filename)) %>% as_tibble()

  ## Load on-board data
  filename = paste0(cruisename, "-X-on-board.csv")
  X_on_board = read.csv(file = file.path(outputdir, filename)) %>% as_tibble()
  
  ## Join them
  joined = full_join(X_cmap, X_on_board, by = "time")
  
  ## Sanity check
  stopifnot(max(abs(joined$lat.x -  joined$lat.y), na.rm = TRUE) < 5E-1)
  stopifnot(max(abs(joined$lon.x -  joined$lon.y), na.rm = TRUE) < 5E-1)
  joined = joined %>% select(-lat.x, -lon.x) %>% rename(lat = lat.y, lon = lon.y)


  ## Make a final visualization of all variables (from joined), 
  ## highlighting the ones from X_on_board as black lines
  df_long <- joined %>% 
    select(-lat, -lon) %>%
    mutate(time = lubridate::as_datetime(time)) %>%
    pivot_longer(-time, names_to = "name", values_to = "value")
  
  hourly_plot <- ggplot(df_long, aes(x = time, y = value, group = name)) +
    geom_line(color = "blue", alpha = 0.2, linewidth = 0.1) +
    geom_line(data = filter(df_long, name %in% onboard_vars), color = "black", linewidth = 0.6) +
    scale_x_datetime(date_breaks = "1 day", labels = scales::label_date("%b %d 00:00")) +
    theme_minimal() +
    theme(panel.grid.minor.x = element_blank()) +
    ggtitle(paste0("All covariates (hourly) over time: ", cruisename)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  # Optional: print or save the plot if needed
  print(hourly_plot)

  ## Make a final visualization of all variables against latitude, connected by time
  df_long <- joined %>% 
    select(-lon) %>%
    mutate(time = lubridate::as_datetime(time)) %>%
    pivot_longer(-c(time, lat), names_to = "name", values_to = "value") %>%
    arrange(name, time) # Crucial: ensures rows are ordered chronologically
   
  lat_plot <- ggplot(df_long, aes(x = lat, y = value, group = name)) +
    geom_path(color = "blue", alpha = 0.2, linewidth = 0.1) + # geom_path respects time order
    geom_path(data = filter(df_long, name %in% onboard_vars), color = "black", linewidth = 0.6) +
    theme_minimal() +
    labs(
      x = "Latitude", 
      y = "Value", 
      title = paste0("All covariates over latitude (connected by time): ", cruisename)
    )
  
  # Print plot
  print(lat_plot)
  
  ## Save to file
  filename = paste0(cruisename, "-X-combined.csv")
  write.csv(joined, file = file.path(outputdir, filename), row.names = FALSE)
}
```





