# Gradients 3 needs some manual processing

First, looking at censor-removed, binned data.

```{r, fig.width = 10, fig.height = 3}
## Load binned data
cruisename = "KM1906"
datadir = "/Users/sangwonh/repos/process-data/output/seaflow/"
filename = paste0(cruisename, "-gridded-without-censor-removal.csv")
list.files(datadir)
hourly_gridded = read.csv(file = file.path(datadir, filename)) %>% as_tibble()
ylist = hourly_gridded %>% group_by(date) %>% group_split() %>%
  lapply(function(df){
    df %>% select(fsc_small_coord, chl_small_coord, pe_coord) %>% as.matrix()
  })
biomass_list = hourly_gridded %>% group_by(date) %>% group_split() %>%
  lapply(function(df){
    df %>% pull(Qc)
  })
dates = lubridate::as_datetime(unique(hourly_gridded$date))

## Plotting before processing
datobj = flowmix::collapse_3d_to_1d(ylist = ylist, countslist = biomass_list, idim = 1)
gg_before = my_plotter(datobj$ylist, datobj$countslist, dates, paste0(cruisename, " before any processing"))


## Find the cutoff's bin number (in fsc_small)
cutoffs = list()
new_countslist = list()
april_16 = which(lubridate::day(dates) == 16)
april_17 = which(lubridate::day(dates) == 17)
april_18 = which(lubridate::day(dates) == 18)
TT = length(biomass_list)
for(tt in 1:TT){

  one_count = datobj$countslist[[tt]]
  one_y = datobj$ylist[[tt]] 

  ## Find cutoff point
  diffs = one_count  %>% diff()
  a = which(sapply(1:(length(diffs)-1), function(ii){
    (sign(diffs[ii]) != sign(diffs[ii+1])) &
      (diff(diffs)[ii]>0)
  }))
  a = min(a, na.rm = TRUE)
  if(tt %in% c(april_16, april_17, april_18)){
    a = a[a<=9] 
  } else {
    a = a[a<=5] 
  }
  stopifnot(length(a) <= 1)
  cutoffs[[tt]] = one_y[a]
}

## Next, removing the points with small fsc_small altogether.
## (Also remove the top-most row in fsc_small and chl_small.)
new_biomass_list = biomass_list
new_ylist = ylist
for(tt in 1:TT){
  one_count = biomass_list[[tt]]
  one_y = ylist[[tt]]
  if( length(cutoffs[[tt]]) > 0 ){
    irow_to_remove1 = which(one_y[,"fsc_small_coord"] <= cutoffs[[tt]])
  } else {
    irow_to_remove1 = c()
  }

  ## Removing the top rows
  irow_to_remove2 = which(one_y[,"fsc_small_coord"] == 80)
  irow_to_remove3 = which(one_y[,"chl_small_coord"] == 40)
  irow_to_remove = c(irow_to_remove1, irow_to_remove2, irow_to_remove3)
  if(length(irow_to_remove) > 0){
  new_ylist[[tt]] = one_y[-irow_to_remove,]
  new_biomass_list[[tt]] = one_count[-irow_to_remove]
  }
}

## Remove these odd time points
## dates[c(21, 136:139)]
## "2019-04-11 19:00:00 UTC"
## "2019-04-16 16:00:00 UTC"
## "2019-04-16 17:00:00 UTC"
## "2019-04-16 18:00:00 UTC"
## "2019-04-16 19:00:00 UTC"
new_dates = dates[-c(21, 136:139)]
new_ylist = new_ylist[-c(21, 136:139)]
new_biomass_list = new_biomass_list[-c(21, 136:139)]
stopifnot(length(new_dates)==length(new_ylist))
stopifnot(length(new_dates)==length(new_biomass_list))

datobj_after_correction = flowmix::collapse_3d_to_1d(ylist = new_ylist, countslist = new_biomass_list, idim = 1)
gg_after = my_plotter(datobj_after_correction$ylist,
                  datobj_after_correction$countslist,
                  new_dates,
                  paste0(cruisename, " after any processing"))
gg_after

cowplot::plot_grid(plotlist = list(gg_before, gg_after), ncol = 1, labels = NULL)    %>% 
  ggsave(filename =  file.path(outputdir, paste0(cruisename, "-processing.png")), width = 10, height = 13)


## Combine these things back into a csv-saveable table.
dates_list <- new_dates %>% as.list()
recombined_list <- purrr::pmap(list(dates_list, new_ylist, new_biomass_list), function(date_str, coords_matrix, qc_vector) {
  coords_tibble <- as_tibble(coords_matrix, .name_repair = "unique")
  colnames(coords_tibble) <- c("fsc_small_coord", "pe_coord", "chl_small_coord") # Rename columns
  coords_tibble %>%
    mutate(date = date_str, Qc = qc_vector)
})
hourly_gridded_reconstructed <- bind_rows(recombined_list) %>%
  select(date, fsc_small_coord, chl_small_coord, pe_coord, Qc)

## Each cruise is now saved to a file
filename = paste0(cruise, "-gridded-after-censor-removal.csv")
outfile = file.path(outputdir, filename)
write.csv(hourly_gridded_reconstructed, file = outfile, row.names = FALSE)
cat("wrote to", outfile, fill=TRUE)
```

# Now, look at G1, G2 and G4

With the same critical pipeline ((1) remove edge points in fsc and chl, (2)
visually remove weird points).

```{r}
## Load binned data
cruises = c("SCOPE_16", ## G1
            "MGL1704", ## G2
            "TN397_740") ## G4
for(cruisename in cruises){
  print(cruisename)
  ## cruisename = "MGL1704"
  ## cruisename = "SCOPE_16"
   cruisename = "TN412_130"
   cruisename = "TN397_740"
  datadir = "/Users/sangwonh/repos/process-data/output/seaflow/"
  filename = paste0(cruisename, "-gridded-without-censor-removal.csv")
  hourly_gridded = read.csv(file = file.path(datadir, filename)) %>% as_tibble()
  hourly_gridded
  ylist = hourly_gridded %>% group_by(date) %>% group_split() %>%
    lapply(function(df){
      df %>% select(fsc_small_coord, chl_small_coord, pe_coord) %>% as.matrix()
    })
  biomass_list = hourly_gridded %>% group_by(date) %>% group_split() %>%
    lapply(function(df){
      df %>% pull(Qc)
    })
  dates = lubridate::as_datetime(unique(hourly_gridded$date))
  
  ## Collapse into 1d for plotting
  datobj = flowmix::collapse_3d_to_1d(ylist = ylist, countslist = biomass_list, idim = 1)
  
  ## Plotting before processing
  gg_before = my_plotter(datobj$ylist, datobj$countslist, dates, paste0(cruisename, " before any processing"))
  
  ## For G2 and G4 (but not G1) remove *top* fsc_small_coord bin altogether
  new_biomass_list = biomass_list
  new_ylist = ylist
  new_dates = dates
  if(cruisename != "SCOPE_16"){
    TT = length(ylist)
    for(tt in 1:TT){
      print(tt)
      one_count = biomass_list[[tt]]
      one_y = ylist[[tt]]
      irow_to_remove2 = which(one_y[,"fsc_small_coord"] == 80)
      irow_to_remove = c(irow_to_remove1, irow_to_remove2)
      if(length(irow_to_remove) > 0){
        new_ylist[[tt]] = one_y[-irow_to_remove,]
        new_biomass_list[[tt]] = one_count[-irow_to_remove]
      }
    }
  }
  
  ## Plotting after processing
  datobj = flowmix::collapse_3d_to_1d(ylist = new_ylist, countslist = new_biomass_list, idim = 1)
  gg_after = my_plotter(datobj$ylist, datobj$countslist, dates, paste0(cruisename, " after any processing"))
  cowplot::plot_grid(plotlist = list(gg_before, gg_after), ncol = 1, labels = NULL)    %>% 
    ggsave(filename =  file.path(outputdir, paste0(cruisename, "-processing.png")), width = 10, height = 13)


  ## Combine these things back into a csv-saveable table.
  dates_list <- new_dates %>% as.list()
  recombined_list <- purrr::pmap(list(dates_list, new_ylist, new_biomass_list), function(date_str, coords_matrix, qc_vector) {
    coords_tibble <- as_tibble(coords_matrix, .name_repair = "unique")
    colnames(coords_tibble) <- c("fsc_small_coord", "pe_coord", "chl_small_coord") # Rename columns
    coords_tibble %>%
      mutate(date = date_str, Qc = qc_vector)
  })
  hourly_gridded_reconstructed <- bind_rows(recombined_list) %>%
    select(date, fsc_small_coord, chl_small_coord, pe_coord, Qc)
  
  ## Each cruise is now saved to a file
  filename = paste0(cruisename, "-gridded-after-censor-removal.csv")
  outfile = file.path(outputdir, filename)
  write.csv(hourly_gridded_reconstructed, file = outfile, row.names = FALSE)
  cat("wrote to", outfile, fill=TRUE)
}
```














Okay, I think cleanup is done, and jobs can be run now. 


```{r}
cruises = c("SCOPE_16", ## G1
            "MGL1704", ## G2
            "TN397_740") ## G4
cruisename = "TN397_740"
datadir = "/Users/sangwonh/repos/process-data/output/seaflow/"
filename = paste0(cruisename, "-gridded-after-censor-removal.csv")
hourly_gridded = read.csv(file = file.path(datadir, filename)) %>% as_tibble()
ylist = hourly_gridded %>% group_by(date) %>% group_split() %>%
  lapply(function(df){
    df %>% select(fsc_small_coord, chl_small_coord, pe_coord) %>% as.matrix()
  })
biomass_list = hourly_gridded %>% group_by(date) %>% group_split() %>%
  lapply(function(df){
    df %>% pull(Qc)
  })
dates = lubridate::as_datetime(unique(hourly_gridded$date))
tt_end = 200
## tt_end = length(dates)
set.seed(0)
obj = flowtrend::flowtrend_once(ylist = ylist[1:tt_end],
                                countslist = biomass_list[1:tt_end],
                                lambda = .01,
                                lambda_prob = .01,
                                numclust = 5,
                                maxdev = 5,
                                l = 1,
                                l_prob = 1, verbose = TRUE,
                                rho_init = 0.05,
                                niter = 10)
## lot(obj$objectives)

?flowtrend::flowtrend_once

datobj = flowmix::collapse_3d_to_1d(ylist = ylist, countslist = biomass_list, idim = 1)
flowtrend::plot_1d(ylist = datobj$ylist[1:tt_end],
                   countslist = datobj$countslist[1:tt_end], bin = TRUE)  +
  geom_line(aes(x=time, y = value, group = name, col = name), data = mnmat_long)

mnmat = obj$mn %>% .[,1,] 
mnmat = as_tibble(mnmat) %>% add_column(time = 1:tt_end) 
mnmat_long = mnmat %>%   pivot_longer(-time)


mnmat
probmat = obj2$prob
matplot(probmat)

```


Things to do next:

1. Fix flowtrend from these errors, methodically
2. Ready the code to submit bioprovince.
3. Read the captions for
