---
title: "Data imputation and visualization for Environmental Factors"
author: "Sangwon Hyun"
date: "2026-08-18"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
execute:
  warning: false
  message: false
---

## 1. Environment Setup & Helper Functions

In this section, we load the necessary R packages and set up matrix visualizers.

```{r setup}
#| label: setup-packages

## Core tidyverse packages
library(tidyverse)
library(ggplot2)
library(tidyr)
library(dplyr)
library(scales)

## Auxiliary packages
library(broom)
library(janitor)
library(softImpute)

## Helper function for plotting matrix heatmaps
drawmat_precise_ggplot <- function(distmat, colours = c("blue", "red"),
                                   limits = NULL,
                                   xlab = "", ylab = "", title = ""){
  distmat <- as.matrix(distmat)
  longData <- reshape2::melt(distmat[nrow(distmat):1, ])

  p <- ggplot() +
    geom_raster(aes(x = Var2, y = Var1, fill = value), data = longData) +
    scale_fill_gradientn(colours = colours, guide = "colorbar", limits = limits) +
    labs(x = xlab, y = ylab, title = title) +
    theme_minimal() + 
    theme(
      axis.text.x = element_text(size = rel(1), angle = 90, vjust = 0.3),
      axis.text.y = element_text(size = rel(1)),
      plot.title = element_text(size = rel(1.5))
    )
  return(p)
}
```

---

## 2. Data Loading & Standardization

Load the cruise dataset and scale numeric features.

```{r load-data}
#| label: load-data

## Load data
X <- read.csv("~/Downloads/X-MGL1704.csv") %>% as_tibble()
print(head(X))

## Notes (8/17/2025):
## (1) PAR resolution is: 0.25 degree vs 1 degree
## (2) Want to make heatmaps.
## (3) Climatology (e.g., WOA) is most dependable; PAR (from CMAP) not so much.

## Extract spatiotemporal identifiers and scale numerical variables
time0 <- lubridate::as_datetime(X$time)
lat0 <- X %>% pull(lat)
lon0 <- X %>% pull(lon)
varnames <- X %>% select(-c("time", "lat", "lon")) %>% colnames()

X_scaled <- X %>%
  select(-lat, -lon, -time) %>%
  scale() %>% 
  as_tibble()

X_combined <- cbind(time = time0, X_scaled)
```

Plot the initial scaled variables across time:

```{r fig-raw-scaled-lines}
#| label: fig-raw-scaled-lines
#| fig-width: 10
#| fig-height: 4
#| fig-cap: "Raw scaled variables over time"

X_combined %>% 
  pivot_longer(-time) %>% 
  ggplot() +
  geom_path(aes(x = time, y = value, group = name), alpha = 0.1, linewidth = 0.1) +
  theme_minimal()
```

---

## 3. Missing Value Diagnostics

Analyze missingness patterns across all variables.

> **Note:** Variables like salinity from GO-SHIP or GEOTRACES trajectories are naturally unavailable for non-targeted cruises ([GO-SHIP Paper](https://www.nature.com/articles/s41597-022-01212-w), [GEOTRACES Overview](https://www.geotraces.org/cruise-overview/)).

```{r fig-missing-histogram}
#| label: fig-missing-histogram
#| fig-width: 8
#| fig-height: 5
#| fig-cap: "Distribution of missing data proportions across variables"

## Proportion of missing time points
num_na <- apply(X_scaled, 2, function(mycol) sum(is.na(mycol))) 
print(head(sort(num_na, decreasing = TRUE)))

num_time_points <- nrow(X_scaled)
prop_missing <- num_na / num_time_points

## Missingness distribution histogram
hist(prop_missing, breaks = 30, main = "Proportion of missing time points")
abline(v = 0.25, col = "red", lty = 2)

missing_only_few <- sum(prop_missing < 0.25)
legend("topleft", legend = paste0("Missing only\na few:\n", missing_only_few, "\nvariables"), box.lwd = 0)

missing_a_lot <- sum(prop_missing > 0.25)
legend("topright", legend = paste0("Missing \na lot:\n", missing_a_lot, "\nvariables"), box.lwd = 0)
```

Visualize the feature matrix before imputation (variables with $<25\%$ missing values):

```{r fig-matrix-before-imputation}
#| label: fig-matrix-before-imputation
#| fig-width: 10
#| fig-height: 6
#| fig-cap: "Matrix heatmap before imputation (<25% missing)"

X_scaled_copy0 <- X_scaled %>% as.matrix()
dimnames(X_scaled_copy0) <- NULL

X_scaled_copy0 %>% 
  .[, prop_missing < 0.25] %>% 
  drawmat_precise_ggplot() + 
  ggtitle("Before Imputation")
```

---

## 4. Matrix Completion via Low-Rank SVD (`softImpute`)

Fill missing values using nuclear-norm regularized singular value decomposition (`softImpute`).

```{r matrix-imputation}
#| label: matrix-imputation

X_scaled_copy <- X_scaled_copy0[, prop_missing < 0.25] 

## Fit nuclear-norm regularized SVD model
fit <- softImpute(X_scaled_copy, rank.max = 100, lambda = 10) ## I haven't carefully chosen lambda (or rank.max); we should do this..

## Complete matrix (fill NAs using low-rank fit)
X_hat <- complete(X_scaled_copy, fit, unscale = FALSE)
```

```{r fig-matrix-after-imputation}
#| label: fig-matrix-after-imputation
#| fig-width: 10
#| fig-height: 6
#| fig-cap: "Matrix heatmap after softImpute low-rank completion"

drawmat_precise_ggplot(X_hat) + ggtitle("After Imputation")
```

### 4.1 Evaluation & Imputed Values Comparison

Compare imputed entries against original observed series:

```{r fig-imputation-facets}
#| label: fig-imputation-facets
#| fig-width: 12
#| fig-height: 8
#| fig-cap: "Comparison of observed vs. softImpute filled values across series"

X_smaller <- X_scaled_copy
observed_vec <- as.vector(X_smaller)
p <- ncol(X_smaller)
n <- nrow(X_smaller)

df <- tibble(
  Time = rep(1:n, times = p),
  Series = rep(1:p, each = n),
  Value = as.vector(X_hat),
  Imputed_or_not = is.na(observed_vec)  
)

df_plot <- df %>%
  mutate(
    Series_Label = paste("Series", Series),
    Status = ifelse(Imputed_or_not, "Imputed", "Observed")
  )

df_plot %>%
  group_by(Series) %>%
  filter(any(Imputed_or_not)) %>%
  ggplot(aes(x = Time, y = Value)) +
  geom_line(color = "gray50", linewidth = 0.5, alpha = 0.7) +
  geom_point(
    data = . %>% filter(Imputed_or_not),
    aes(color = Status),
    size = 0.3
  ) +
  facet_wrap(~ Series_Label, scales = "free_y") +
  scale_color_manual(
    values = c("Imputed" = "firebrick", "Observed" = "gray30")
  ) +
  labs(
    title = "Time Series Imputation via softImpute",
    x = "Time Index",
    y = "Value",
    color = "Data Point"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold")
  )
```

Confirm non-imputed entries were untouched:

```{r sanity-check}
#| label: sanity-check

(X_scaled_copy0[, prop_missing < 0.25] - X_hat) %>% 
  as.numeric() %>% 
  summary()
```

---

## 5. Hourly Aggregation & Spatiotemporal Plots

Format the imputed output, re-attach spatial coordinates, and aggregate to hourly resolution.

```{r hourly-aggregation}
#| label: hourly-aggregation

colnames(X_hat) <- varnames[prop_missing < 0.25]

X_imputed <- X_hat %>% 
  as_tibble() %>% 
  add_column(time = time0, lat = lat0, lon = lon0, .before = 1)

## Aggregate by hour using floor_date
X_imputed_hourly <- X_imputed %>% 
  group_by(time = lubridate::floor_date(time, "hour")) %>%
  summarize(across(where(is.numeric), mean), .groups = "drop")
```

```{r fig-imputed-hourly-time}
#| label: fig-imputed-hourly-time
#| fig-width: 10
#| fig-height: 4
#| fig-cap: "Hourly aggregated variables over time (00:00 daily ticks)"

X_imputed_hourly %>% 
  select(-lat, -lon) %>%
  pivot_longer(-time) %>%
  ggplot() +
  geom_line(aes(x = time, y = value, group = name), col = "blue", alpha = 0.2, linewidth = 0.1) +
  scale_x_datetime(
    breaks = "1 day",
    labels = label_date("%b %d\n00:00") 
  ) +
  theme_minimal() +
  theme(panel.grid.minor.x = element_blank())
```

```{r fig-imputed-hourly-latitude}
#| label: fig-imputed-hourly-latitude
#| fig-width: 10
#| fig-height: 4
#| fig-cap: "Hourly aggregated variables plotted across latitude"

X_imputed_hourly %>% 
  pivot_longer(-c(time, lat, lon)) %>% 
  ggplot(aes(x = lat, y = value, group = name)) +
  geom_path(col = "blue", linewidth = 0.1, alpha = 0.2) +
  theme_minimal()
```

---

## 6. Principal Component Analysis (PCA)

Run PCA on observation time points across all imputed variables.

```{r pca-analysis}
#| label: pca-analysis

pca_res <- X_imputed_hourly %>% 
  select(-time, -lat, -lon) %>% 
  prcomp(scale. = TRUE) %>%
  broom::augment(X_imputed_hourly %>% select(time, lat, lon)) %>% 
  janitor::clean_names() %>% 
  rename_with(~ gsub("^fitted_", "", .x), starts_with("fitted_")) %>%
  select(time, lon, lat, matches("^pc[1-6]$"))
```

```{r fig-pca-components}
#| label: fig-pca-components
#| fig-width: 10
#| fig-height: 6
#| fig-cap: "First six principal components across time"

pca_res %>% 
  select(-lat, -lon) %>% 
  pivot_longer(-time) %>%
  ggplot() +
  geom_line(aes(x = time, y = value, group = name)) +
  facet_wrap(~ name) +
  theme_minimal()
```

---

## 7. Next Directions

1. Integrate onboard lab variables (PAR, salinity, temperature).
2. Perform unsupervised variable space embeddings (PCA/UMAP on transposed feature matrix).
