---
title: "Data imputation and visualization for Environmental Factors"
author: "Sangwon Hyun"
date: "2026-08-18"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
execute:
  warning: false
  message: false
---

## 1. Environment Setup & Helper Functions

In this section, we load the necessary R packages and set up matrix visualizers.

```{r setup}
#| label: setup-packages

## Core tidyverse packages
library(tidyverse)
library(ggplot2)
library(tidyr)
library(dplyr)
library(scales)

## Auxiliary packages
library(broom)
library(janitor)
library(softImpute)

## Helper function for plotting matrix heatmaps
drawmat_precise_ggplot <- function(distmat, colours = c("blue", "red"),
                                   limits = NULL,
                                   xlab = "", ylab = "", title = ""){
  distmat <- as.matrix(distmat)
  longData <- reshape2::melt(distmat[nrow(distmat):1, ])

  p <- ggplot() +
    geom_raster(aes(x = Var2, y = Var1, fill = value), data = longData) +
    scale_fill_gradientn(colours = colours, guide = "colorbar", limits = limits) +
    labs(x = xlab, y = ylab, title = title) +
    theme_minimal() + 
    theme(
      axis.text.x = element_text(size = rel(1), angle = 90, vjust = 0.3),
      axis.text.y = element_text(size = rel(1)),
      plot.title = element_text(size = rel(1.5))
    )
  return(p)
}
```

---

## 2. Data Loading & Standardization

Also load the handful of environmental covariates from on-board the cruise.


```{r}
raw_data_dir <- "~/Dropbox/data/ocean/raw-seaflow/particle-data-2025-11-21"
list.files(raw_data_dir)
cruises = c("SCOPE_16", ## G1
            "MGL1704", ## G2
            "KM1906", ## G3
            "TN397_740", ## G4
            "TN412_130") ## G5
datlist = lapply(cruises, function(cruisename){

  
  ## ## temporary
  ## cruisename = "MGL1704"

  ## Load sunlight (par), temperature, conductivity, salinity
  geodat =
    file.path(raw_data_dir, cruisename, "dash", paste0(cruisename, "-geo.tsdata"))   %>%
    read.csv(sep = "\t", skip = 6)  %>% as_tibble() %>%
    mutate(time = lubridate::as_datetime(time))
  
  geodat =
    geodat %>%
    mutate(hour_group = floor_date(time, "1 hour")) %>%
    group_by(hour_group) %>%
    summarise(across(where(is.numeric), \(x) mean(x, na.rm=TRUE))) %>%
    rename(time = hour_group)
  
  ## Manual adjustments
  if(cruisename == "MGL1704"){


    geodat = geodat %>%
      mutate(
          ## 1. Turn the bad value into NA
          salinity = if_else(time %in% 
                             c(lubridate::as_datetime("2017-06-05 01:00:00"),
                               lubridate::as_datetime("2017-06-05 02:00:00"),
                               lubridate::as_datetime("2017-06-05 03:00:00"),
                               lubridate::as_datetime("2017-06-05 04:00:00")), NA_real_, salinity),
    # 2. Automatically interpolate missing values based on the time vector
    salinity = zoo::na.approx(salinity, x = time, na.rm = FALSE)
  )

    geodat = geodat %>%
      mutate(
          ## 1. Turn the bad value into NA
          conductivity = if_else(time %in% 
                             c(lubridate::as_datetime("2017-06-05 01:00:00"),
                               lubridate::as_datetime("2017-06-05 02:00:00"),
                               lubridate::as_datetime("2017-06-05 03:00:00"),
                               lubridate::as_datetime("2017-06-05 04:00:00")), NA_real_, conductivity),
    # 2. Automatically interpolate missing values based on the time vector
    conductivity = zoo::na.approx(conductivity, x = time, na.rm = FALSE)
  )

  }

  if(cruisename == "SCOPE_16"){
    ## Clean one single point using the average of the +-1 hours
    t_target <- lubridate::as_datetime("2016-05-02 05:00:00")
    t_prev   <- lubridate::as_datetime("2016-05-02 04:00:00")
    t_next   <- lubridate::as_datetime("2016-05-02 06:00:00")
    geodat$salinity[geodat$time == t_target] <- 
      (geodat$salinity[geodat$time == t_prev] + geodat$salinity[geodat$time == t_next]) / 2
  }

  ## Manual adjustments
  if(cruisename == "KM1906"){
    geodat = geodat %>% filter(time > lubridate::as_datetime("2019-04-10 12:00:00"))  

    ## Clean one single point using the average of the +-1 hours
    t_target <- lubridate::as_datetime("2019-04-15 03:00:00")
    t_prev   <- lubridate::as_datetime("2019-04-15 02:00:00")
    t_next   <- lubridate::as_datetime("2019-04-15 04:00:00")
    geodat$salinity[geodat$time == t_target] <- 
      (geodat$salinity[geodat$time == t_prev] + geodat$salinity[geodat$time == t_next]) / 2

    ## Impute PAR

  }
  return(geodat)
})
names(datlist) = cruises

plotlist <- lapply(cruises, function(cruisename){
  geodat = datlist[[cruisename]]
  geodat %>%
    pivot_longer(-time) %>%
    mutate(name = factor(name, levels = c("lat", "lon", "temp", "par", "salinity", "conductivity"))) %>% 
    ggplot() +
    geom_line(aes(x=time, y = value, group = name, col = name)) +
    facet_wrap(~name, scale = "free_y", ncol = 6) +
    ggtitle(cruisename)
    })
cowplot::plot_grid(plotlist = plotlist, nrow=5)


lapply(names(datlist), function(nm) {
  dat <- datlist[[nm]]
  drawmat_precise_ggplot(dat %>% select(-c("time")) %>% as.matrix()%>% scale()) +
    ggtitle(nm)
}) %>%
  cowplot::plot_grid(plotlist = ., ncol = 5)


combined_dat = datlist %>% bind_rows(.id = "cruisename") %>% rename(temp_onboard = temp,
                                                                    salinity_onboard = salinity,
                                                                    par_onboard = par,
                                                                    conductivity_onboard = conductivity)
```

## 2. Combine with CMAP-downloaded covariates

```{r}

## (list(X_before_scaling = X,
##              X_after_scaling = X_scaled,
##              scaleobj = scaleobj,
##              X_means = mean_val,
##              X_sds = sd_val),
cruisename = "MGL1704"
filename = paste0(cruisename, "-scaling-info.RDS")
covariates_from_cmap <- readRDS(file = file.path("~/repos/process-data/output/seaflow", filename))

```


Next up
- [ ] Combine these with `X.RDS`.
- [ ] Send them to Trisha.
- [ ] Dig up the old covariates (where are they..).
- [ ] Fit flowmix/flowtrend on these models.
- [ ] Make a plan for doing the rest.

